---
title: "TRY allometric models"
author: "Marcos Longo"
date: "2023-01-09"
output: html_document
---

```{=html}
<style>
pre {
  overflow-x: auto;
}
pre code {
  word-wrap: normal;
  white-space: pre;
}
</style>
```
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction.

This script seeks to establish allometric relationships for different plant functional groups.

Set paths and file names for input and output.

-   **home_path**. The user home path.
-   **main_path**. Main working directory
-   **util_path**. The path with the additional utility scripts (the full path of `RUtils`).
-   **summ_path**. Path for the allometry summary.
-   **plot_path**. The main output path for the simulation plots.
-   **rdata_path**. Output path for R objects (so we can use it for comparisons.)

```{r, label='path-settings',message=FALSE,results='hide',warning=FALSE}
# Set useful paths and file names
home_path       = path.expand("~")
main_path       = file.path(home_path,"Data", "TraitAllom_Workflow")
util_path       = file.path(main_path,"RUtils")
summ_path       = file.path(main_path,"TaxonSummary")
plot_path       = file.path(main_path,"Figures")
rdata_path      = file.path(main_path,"RData")
```

The following block defines some settings that were used for defining clusters and standardised major axes. This is used to reload the correct data sets and to generate unique labels for the allometry objects.

-   **use_lifeclass**. Which life-form/phylogenetic level was used to subset the original data set. Options are:

    -   `"FlowerTrees"`. Trees, classes Magnoliopsida and Liliopsida (flowering plants).
    -   `"Shrubs"`. Shrubs
    -   `"Grasses"`. Grasses/Herbs
    -   `"FlowerPlants"`. All life forms, classes Magnoliopsida and Liliopsida
    -   `"Pinopsida"`. Conifers (class Pinopsida), all life forms
    -   `"SeedPlants"`. Seed plants, all life forms: classes Cycadopsida, Ginkgoopsida, Gnetopsida, Liliopsida, Magnoliopsida and Pinopsida.
    -   `"Plantae"`. All plants

-   **use_realm**. Which realm was used to subset the original data set. Current options are:

    -   `"NeoTropical"`. South and Central America
    -   `"PanTropical"`. All continents.
    -   `"AfricaTropical"`. Africa
    -   `"AsiaTropical"`. Asia
    -   `"WestUS"`. Western US

```{r, label='set-sma-options',message=FALSE,results='hide',warning=FALSE}

# Life-form/phylogenetic level to use for SMA analyses.
use_lifeclass  = c("FlowerTrees","Shrubs","Grasses","FlowerPlants","Pinopsida","SeedPlants","Plantae")[4L]

# Realm to use for SMA analyses.
use_realm  = c("NeoTropical","PanTropical", "AfricaTropical", "AsiaTropical", "WestUS")[1L]

# Taxonomic level of detail for SMA analyses. This is used to retrieve the correct "Tidy TRY" data set
# with the cluster assignment. Allometric models are always developed from individual observations.
fit_taxon = c("Individual","Species","Genus")[2L]
```



Settings for reloading or rerunning multiple steps of this script. These are all logical variables, with `TRUE` meaning to reload previous calculations, and `FALSE` meaning to calculate the step again. If the RData file is not found, these variables will be ignored.

-  **reload_input_allom**. Reload the input data for fitting the allometric models?
-  **reload_height_allom**. Reload the height-DBH models?
-  **reload_agb_allom**. Reload above-ground biomass models?
-  **reload_bleaf_allom**. Reload leaf biomass models?
-  **reload_leaf_area_allom**. Reload individual leaf area models?
-  **reload_crown_area_allom**. Reload individual crown area models?
-  **reload_crown_depth_allom**. Reload crown depth models?
-  **reload_bark_thick_allom**. Reload bark thickness models?
```{r, label='set-reload',message=FALSE,results='hide',warning=FALSE}
reload_input_allom       = c(FALSE,TRUE)[2L] # Reload the input data for fitting the allometric models?
reload_height_allom      = c(FALSE,TRUE)[2L] # Reload the height-DBH models?
reload_agb_allom         = c(FALSE,TRUE)[2L] # Reload above-ground biomass models?
reload_bleaf_allom       = c(FALSE,TRUE)[2L] # Reload leaf biomass models?
reload_leaf_area_allom   = c(FALSE,TRUE)[2L] # Reload individual leaf area models?
reload_crown_area_allom  = c(FALSE,TRUE)[2L] # Reload individual crown area models?
reload_crown_depth_allom = c(FALSE,TRUE)[2L] # Reload crown depth models?
reload_bark_thick_allom  = c(FALSE,TRUE)[2L] # Reload bark thickness models?
```



The following block defines some settings for model fitting.

-  **CntAllomMin**. Minimum number of valid points to consider for maps.
-  **CntAllomMax**. Maximum number of points to consider for fitting. This is used to limit the model fitting to a more manageable number of observations when allometric data are exceedingly large. If no cap is sought, set `CntAllomMax = +Inf`, which will ensure to always use the actual number of observations. This setting is independent on the data binning sampling described in the next chunk.
-  **AllomConfInt**. Confidence interval for allometric equations.
-  **AllomFitXIC**. Which information criterion to use when comparing models? Options are `"AIC"` (default) for Akaike Information Criterion, `"BIC"` for Bayesian Information Criterion, `"ARMSE"` for the adjusted root mean square error (accounting for the number of
parameters) and `"mR2Adj"` for the negative adjusted coefficient of determination ($-R^2_\textrm{adj}$). The information criteria approaches will pick the model with the highest likelihood (corrected for additional parameters), whereas the `"aRMSE"` and `"wRMSE"` will heavily penalise biased models. As a consequence, information criteria approaches will be more likely to select heteroscedastic or log-normal models, and `"aRMSE"` will likely pick homoscedastic models. Option
`"wRMSE"` is intended to account for both, but it is still experimental and can get the worst of both worlds instead of the best of both worlds.
-  **AllomMaxItOptim**. Maximum number of iterations before giving up optimising the model. This is actually what controls the search for the optimal point, and should be relatively high (1000--5000) as sometimes the initial guess is too far from the actual optimal.
-  **AllomMaxItGain**. Maximum number of iterations for calling the main optimiser for seeking a better maximum using the previous optimised set of parameters. This sometimes helps improving the model, but it typically stabilises after a handful of iterations (about 5).
-  **AllomTolOptim**. Relative tolerance for model fitting. This controls the error tolerance of the optimiser itself and should be stricter ($10^{-7}$) or less).
-  **AllomTolGain**. Relative tolerance for successive calls of the main optimiser. Values of the order of $10^{-5}$ are typically sufficient.
-  **UseFixedModel**. If `TRUE`, use global observations to define the best model formulation (equation type), and use this one for all classes and categories. If `FALSE`, use the test the best formulation for each class and category within class.
-  **AllomCntPred**. Number of points spanning the range of predictors for fitted curve.
-  **AllomQuantPred**. Logical flag: should the values span evenly by quantiles (TRUE) or just linearly (FALSE).
```{r, label='set-map-options',message=FALSE,results='hide',warning=FALSE}
CntAllomMin   = 100L      # Minimum number of points for fitting allometric models.
CntAllomMax   = Inf       # Maximum number of points for fitting allometric models (if more data are available, we sample the training data sets).
AllomConfInt  = 0.95      # Confidence interval for allometric model fittings.

# Information criterion for selecting the best model
AllomFitXIC     = c("AIC","BIC","aRMSE","wRMSE")[1L]

# Settings for non-linear heteroscedastic model fitting.
AllomMaxItOptim  = 10000L   # Max. # of iterations for the gnls optimisation
AllomMaxItGain   = 5L       # Max. # of iterations for the nls optimisation step 
AllomTolOptim    = 1.e-8    # Tolerance for convergence in the gnls algorithm
AllomTolGain     = 1.e-4    # Tolerance for convergence criterion in nls step



# Settings for predicting values.
UseFixedModel  = TRUE     # Use the same functional form (model) for each class category?
AllomCntPred   = 101L     # Number of points spanning the range of predictor for fitted curve.
AllomQuantPred = TRUE     # Span predictor evenly across quantiles (TRUE) or linearly (FALSE)
AllomCntBoot   = 16L      # 1024L    # Number of times for bootstrapping for generating confidence bands.
```


Random number seeds for the multiple models. Set these to specific numbers to ensure code reproducibility, or set them to `NA_integer_` for completely random data points.
- **rseed_input**. Random seed for gap filling traits (typically SLA and wood density) for the allometry training data.
- **rseed_height**. Random seed for fitting the allometry model for height.
- **rseed_agb**. Random seed for fitting the allometry model for above-ground biomass.
- **rseed_bleaf**. Random seed for fitting the allometry model for leaf biomass.
- **rseed_leaf_area**. Random seed for fitting the allometry model for leaf area.
- **rseed_crown_area**. Random seed for fitting the allometry model for crown area.
- **rseed_crown_depth**. Random seed for fitting the allometry model for crown depth.
- **rseed_bark_thick**. Random seed for fitting the allometry model for bark thickness.

```{r, label='set-random-seeds',message=FALSE,results='hide',warning=FALSE}
rseed_input       = 0L   # Random seed - input data gap filling
rseed_height      = 6L   # Random seed - height allometry
rseed_agb         = 12L  # Random seed - above-ground biomass
rseed_bleaf       = 18L  # Random seed - leaf biomass
rseed_leaf_area   = 24L  # Random seed - leaf area
rseed_crown_area  = 30L  # Random seed - crown area
rseed_crown_depth = 36L  # Random seed - crown depth
rseed_bark_thick  = 42L  # Random seed - bark thickness

```

Sampling parameters for the data binning.

- **UseSizeBins**. Should the training date be resampled by size bins? Provide one value (`TRUE`|`FALSE`) for each predicting variable.
- **MinSmpPerBin**. Minimum number of observations for each bin. Binning will not occur unless there are at least twice as many valid points (and possibly even more points in case of very imbalanced data sets).
- **MaxCntBins**. Maximum number of bins to be considered.

```{r, label='set-binning-options',message=FALSE,results='hide',warning=FALSE}
UseSizeBins  = c(height=TRUE,agb=FALSE,bleaf=FALSE,leaf_area=FALSE,crown_area=TRUE,crown_depth=TRUE,bark_thick=FALSE)
MinSmpPerBin = 100L
MaxCntBins   = 50L
```


Set up some flags to either plot (`TRUE`) or skip (`FALSE`) groups of plots.

- **plot_allom_scatter**. Plot scatter plot comparing observations with predicted values?
- **plot_allom_fun**. Plot functional form of models as function of size? 

```{r,label='plot-switches',message=FALSE,results='hide',warning=FALSE}
plot_allom_scatter     = c(FALSE,TRUE)[2L] # Plot allometry scatter plot ? (TRUE|FALSE)
plot_allom_fun         = c(FALSE,TRUE)[2L] # Plot allometry function?      (TRUE|FALSE)
```


General plot options for `ggplot`.

```{r, label='set-ggplot',message=FALSE,results='hide',warning=FALSE}
gg_device        = c("pdf") # Output devices to use (Check ggsave for acceptable formats)
gg_depth         = 300      # Plot resolution (dpi)
gg_ptsz          = 24       # Font size
gg_width         = 11.0     # Plot width for non-map plots (units below)
gg_height        = 8.5      # Plot height for non-map plots (units below)
gg_units         = "in"     # Units for plot size
gg_screen        = TRUE     # Show plots on screen as well?
gg_tfmt          = "%Y"     # Format for time 
gg_ncolours      = 129      # Number of node colours for heat maps.
gg_fleg          = 1./6.    # Fraction of plotting area dedicated for legend
ndevice = length(gg_device)
```


Options for allometry plots:

- **CntDensThresh**. Number of points above which we select density plots instead of scatter plots. Note: if the plot uses logarithmic scale in either axis, the code will force scatter plots.
- **CntDensBin**. Number of bins for 2-D density plots.
- **DensPalette**. Colour palette for density bins.
- **DensReverse**. Should we use the reverse palette? (TRUE | FALSE)
- **DensCntColour**. Number of colours to use in the colour ramp.
- **DensTrans**. Variable transformation for density plots.
- **DensRange**. Range for the binned plots, in case a fixed range is sought. This is a vector of 2 (minimum and maximum). If you want to use the default values, set them to `NA_real_`. For completely unbounded values, set both values to `NA_real_`.
- **PointColour**. Point colour when using scatter plots instead of density.
- **PointShape**. Point shape when using scatter plots instead of density.
- **PointSize**. Point size when using scatter plots instead of density.

Scales for x and y axes. Options are `"identity"` (linear), `"log"` (logarithmic, useful for non-linear ranges that are always positive), `"neglog"` (negative logarithm of negative number, i.e. $-\log{\left(-x\right)}$, useful for non-linear scales that are always negative), `"sqrt"` (square root, useful for non-linear ranges that must include zero) and `"cbrt"` (cube root, useful for non-linear ranges that can be positive or negative).

- **xTransAllom**. Vector with transformation for the x axis, one for each X predictor (DBH, Height, Size, WDSize, LMSize)
- **xTransAllom**. Vector with transformation for the y axis, one for each allometric relationship (Height, AGB, BLeaf, LeafArea, CrownArea, CrownDepth, BarkThick)

```{r, label='set-allom-plots',message=FALSE,results='hide',warning=FALSE}
CntDensThresh = 1000L
CntDensBin    = 40L
DensPalette   = "Blues"
DensReverse   = FALSE
DensCntColour = 200L
DensTrans     = "log10"
DensRange     = c(1L,NA_real_)
PointColour   = "#418CC3" # Colour for points (when not using density).
PointShape    = 16L      # Point shape format (when not using density).
PointSize     = 0.75      # Point size (when not using density).

xTransAllom   = c( DBH        = "identity", Height     = "identity", Size       = "log"
                 , WDSize     = "log"     , LMSize     = "log"     )
yTransAllom   = c( Height     = "identity", AGB        = "log"     , BLeaf      = "log"
                 , LeafArea   = "log"     , CrownArea  = "log"     , CrownDepth = "identity"
                 , BarkThick  = "log"     )
```

<!-- =================================================================================== -->
<!-- =================================================================================== -->
<!-- =================================================================================== -->
<!-- =================================================================================== -->
<!-- =================================================================================== -->
<!--                                                                                     -->
<!--               CHANGES BEYOND THIS POINT ARE ONLY FOR CODE DEVELOPMENT               -->
<!--                                                                                     -->
<!-- =================================================================================== -->
<!-- =================================================================================== -->
<!-- =================================================================================== -->
<!-- =================================================================================== -->
<!-- =================================================================================== -->

# Main script

**Note:** Code changes beyond this point are only needed if you are developing the notebook.

## Initial settings.

First, we load the list of host land model variables that we may consider for the comparison.

```{r, label='load-everything',message=FALSE,results='hide',warning=FALSE}
source(file.path(util_path,"load.everything.r"),chdir=TRUE)
```


We also list all colour palettes from packages `RColorBrewer` and `viridis`.  This will be useful when deciding the colour ramps.
```{r, label='retrieve-colour-ramps',message=FALSE, results='hide'}
# List of palettes in package RColorBrewer and viridis
BrewerPalInfo  = rownames(RColorBrewer::brewer.pal.info)
ViridisPalInfo = as.character(lsf.str("package:viridis"))
ViridisPalInfo = ViridisPalInfo[! grepl(pattern="^scale_",x=ViridisPalInfo)]
```

Load the harmonised trait data set, stored in file `rdata_TidyTRY`. This is the output of script `TidyTRY.Rmd` so make sure to run that pre-processing first.

```{r, label='tidy-loader',message=FALSE,results='hide',warning=FALSE}
# Set subset of TRY entries.
use_suffix   = paste(use_realm,use_lifeclass         ,sep="_")

# File name with the tidy data set of individual trait observations.
rdata_TidyTRY = file.path(rdata_path,paste0("TidyTRY_" ,use_suffix,".RData")) 

# Check that the file exists and load it.
if (file.exists(rdata_TidyTRY)){
   # Load data.
   cat0(" + Load data from file: ",basename(rdata_TidyTRY),".")
   dummy = load(rdata_TidyTRY)
}else{
   # File not found, stop the script
   cat0(" + File ",basename(rdata_TidyTRY)," not found!")
   cat0("   This script requires pre-processing and subsetting TRY observations.")
   cat0(" - Run script \"TidyTRY.Rmd\" before running this script, and set: ")
   cat0("      use_realm     = \"",use_realm    ,"\"")
   cat0("      use_lifeclass = \"",use_lifeclass,"\"")
   cat0("   in the TidyTRY preamble.")
   stop(" RData object not found.")
}#end if (file.exists(rdata_TidyTRY))


# Build suffix for model fittings.
base_suffix      = paste(use_suffix,fit_taxon,sep="_")
rdata_TidyCluster = file.path(rdata_path,paste0("TidyCluster_",base_suffix,".RData"))

# Check that the file exists and load it.
if (file.exists(rdata_TidyCluster)){
   # Load data.
   cat0(" + Load data from file: ",basename(rdata_TidyCluster),".")
   dummy = load(rdata_TidyCluster)
}else{
   # File not found, stop the script
   cat0(" + File ",basename(rdata_TidyCluster)," not found!")
   cat0("   This script requires pre-processing, subsetting TRY observations, and a cluster analysis.")
   cat0(" - Run scripts \"TidyTRY.Rmd\" and \"TRYTradeOffs.Rmd\" in this orde before running this script, and set: ")
   cat0("      use_realm     = \"",use_realm    ,"\"")
   cat0("      use_lifeclass = \"",use_lifeclass,"\"")
   cat0("   in the \"TidyTRY.Rmd\" and \"TRYTradeOffs.Rmd\" preambles.")
   stop(" RData object not found.")
}#end if (file.exists(rdata_TidyCluster))
```

Define files and paths for input and output. We also create the output paths.

```{r, label='path-define',message=FALSE,results='hide',warning=FALSE}
# Build output directory for trait, allometry, and photosynthesis fits.
allom_path  = file.path(plot_path,paste0("Allom_",base_suffix))

# Build RData object file names for the allometric models.
rdata_input_allom       = file.path(rdata_path,paste0("AllomInput"         ,use_suffix,".RData"))
rdata_height_allom      = file.path(rdata_path,paste0("AllomFit_Height"    ,use_suffix,".RData"))
rdata_agb_allom         = file.path(rdata_path,paste0("AllomFit_AGB"       ,use_suffix,".RData"))
rdata_bleaf_allom       = file.path(rdata_path,paste0("AllomFit_BLeaf"     ,use_suffix,".RData"))
rdata_leaf_area_allom   = file.path(rdata_path,paste0("AllomFit_LeafArea"  ,use_suffix,".RData"))
rdata_crown_area_allom  = file.path(rdata_path,paste0("AllomFit_CrownArea" ,use_suffix,".RData")) 
rdata_crown_depth_allom = file.path(rdata_path,paste0("AllomFit_CrownDepth",use_suffix,".RData"))
rdata_bark_thick_allom  = file.path(rdata_path,paste0("AllomFit_BarkThick" ,use_suffix,".RData"))


# Make sure directories are set.
dummy = dir.create(path=rdata_path     ,showWarnings=FALSE,recursive=TRUE)
dummy = dir.create(path=allom_path     ,showWarnings=FALSE,recursive=TRUE)
```

Define the labels for titles:

```{r,label='set-subtitle-labels',message=FALSE,results='hide',warning=FALSE}
# Label for life-form/phylogenetic level
LabelLife = switch( use_lifeclass
                  , FlowerTrees  = "flowering trees"
                  , Shrubs       = "shrubs"
                  , Grasses      = "grasses"
                  , FlowerPlants = "flowering plants"
                  , Pinopsida    = "conifers"
                  , SeedPlants   = "seed plants"
                  , Plantae      = "plants"
                  , stop("Unrecognised life-form/phylogenetic level.")
                  )#end switch

# Label for floristic realm
LabelRealm = switch( use_realm
                   , NeoTropical = "Neotropical"
                   , PanTropical = "Pantropical"
                   , stop("Unrecognised realm.")
                   )#end switch

# Build sub-title
LabelSubtitle = paste0(LabelRealm," ",LabelLife)
```

Find confidence quantiles based on the confidence range.

```{r, label='set-confidence-band'}
# Find lower and upper confidence bands
AllomConfLwr = 0.5 - 0.5 * AllomConfInt
AllomConfUpr = 0.5 + 0.5 * AllomConfInt

# Find the significance level to consider allometric models significant
AllomAlpha = 1. - AllomConfInt
```

For some plots, we use a square image so it looks nicer. Define the size as the average between height and width.
```{r,label='set-gg-square',message=FALSE,results='hide',warning=FALSE}
# Find size for square plots
gg_square = sqrt(gg_width*gg_height)
```

Define the tibble that helps fitting and plotting allometric models by category

```{r,label='set-categ-allom',message=FALSE,results='hide',warning=FALSE}
# Find the number of sub-classes to test the model
CategTmp  = CategExtra %>%
   filter( ( (XYUse %in% "Colour") | (TraitID %in% 0L) ) & (! duplicated(Class))) %>%
   mutate( TraitName = ifelse( test = TraitID %in% 0L
                             , yes  = "Cluster"
                             , no   = try_trait$Name[match(TraitID,try_trait$TraitID)]
                             )#end ifelse
         )#end mutate
CategAll   = CategExtra %>%
   filter( is.na(TraitID) & (Class %in% "ALL"))

CategAllom = tibble( TraitID    = c(CategAll$TraitID,CategTmp$TraitID)
                   , Name       = c(CategAll$Class,CategTmp$Class)
                   , TraitClass = ifelse( test = TraitID %gt% 0L
                                        , yes  = try_trait$Name[match(TraitID,try_trait$TraitID)]
                                        , no   = ifelse(test=TraitID %eq% 0L, yes="Cluster",no="All")
                                        )#end ifelse
                   , DescClass  = ifelse( test = TraitID %gt% 0L
                                        , yes  = try_trait$Desc[match(TraitID,try_trait$TraitID)]
                                        , no   = ifelse( test = TraitID %eq% 0L
                                                       , yes  = "Data-based cluster"
                                                       , no   = "All data"
                                                       )#end ifelse
                                        )#end ifelse
                   , Colour     = c(CategAll$Colour,CategTmp$Colour)
                   , Symbol     = c(CategAll$Symbol,CategTmp$Symbol)
                   , XYUse      = c(CategAll$XYUse ,CategTmp$XYUse )
                   )#end tibble

CntCategAllom = nrow(CategAllom)
```


# Prepare data for model fitting

Here we will generate tibble `AllomTRY`, which will be the based on `TidyTRY`, but with the following differences:

1. Wood density ($\rho_{w}$) and specific leaf area ($\mathrm{SLA}$), when available, will be gap filled. If the species has a value associated with the trait, we use that value, otherwise, we randomly sample from either the cluster or using the global distribution of values.
2. New variable leaf mass per area ($\mathrm{LMA}$) as the reciprocal of $\mathrm{SLA}$. This may be useful for defining allometric models.
3. We will define three new variables that are derived from diameter at breast height ($\mathrm{DBH}$) and plant height ($h$), as well as $\rho_w$ or $\mathrm{LMA}$ which are often used in allometric models.
  a. Size - This is defined as $\mathrm{DBH}^2 \, h$
  b. WDSize - This is defined as $\rho_w \, \mathrm{Size}$
  b. LMSize - This is defined as $\mathrm{LMA} \, \mathrm{Size}$

```{r, label='try-prepare-allom',message=FALSE,results='hide',warning=FALSE}
if (reload_input_allom && file.exists(rdata_input_allom)){
   # Load existing data.
   cat0(" + Load input allometry data from ",basename(rdata_input_allom),".")
   dummy = load(rdata_input_allom)
}else{
   # Set random seed
   if (is.na(rseed_input)){
      SeedPrep    = Sys.time()
      rseed_input = 3600*hour(SeedPrep) + 60*minute(SeedPred) + floor(second(SeedPrep))
   }#end if (is.na(rseed_input))
   dummy = set.seed(rseed_input)

   cat0(" + Create input data set for training allometric models.")
   # Create tibble AllomTRY based on TidyTRY
   AllomTRY = TidyTRY
   

   cat0(" + Create template for appending new traits as needed.")
   try_template = try_trait[1L,,drop=FALSE] %>%
      mutate( across(where(is.character), ~ NA_character_)
            , across(where(is.integer  ), ~ NA_integer_  )
            , across(where(is.double   ), ~ NA_real_     )
            , across(where(is.logical  ), ~ FALSE        )
            , across(matches("Type")    , ~ "numeric"    )
            , across(matches("Add" )    , ~ 0.           )
            , across(matches("Mult" )   , ~ 1.           )
            , across(matches("Power")   , ~ 1.           ) )

   
   #    Make sure that the main predictors and predictands are all defined. If they are not
   # part of the original data set, add dummy vectors.
   TraitCheck = tribble( ~Name       , ~IDList                     , ~Desc                          , ~Unit   , ~Trans
                       , "SLA"       , "c(3086L,3115L,3116L,3117L)", "Specific Leaf Area"           , "m2okg" , "log"
                       , "WoodDens"  ,                      "c(4L)", "Wood Density"                 , "gocm3" , "identity"
                       , "DBH"       ,                     "c(21L)", "Diameter at Breast Height"    , "cm"    , "identity"
                       , "Height"    ,               "c(18L,3106L)", "Plant Height"                 , "m"     , "identity"
                       , "AGB"       ,                   "c(3446L)", "Aboveground Biomass"          , "kg"    , "log"
                       , "BLeaf"     ,              "c(129L,3441L)", "Leaf Biomass"                 , "kg"    , "log"
                       , "LeafArea"  ,                    "c(410L)", "Leaf Area per Plant"          , "m2"    , "log"
                       , "CrownArea" ,                     "c(20L)", "Crown Area per Plant"         , "m2"    , "log"
                       , "CrownDepth",                    "c(773L)", "Crown Depth"                  , "m"     , "log"
                       , "BarkThick" ,                     "c(24L)", "Bark thickness"               , "cm"    , "identity"
                       , "LMA"       ,                     "c(-1L)", "Leaf Mass per Area"           , "kgom2" , "log"
                       , "Size"      ,                     "c(-2L)", "Tree size"                    , "cm2m"  , "log"
                       , "WDSize"    ,                     "c(-3L)", "Wood-Density Scaled Tree Size", "cg"    , "log"
                       , "LMSize"    ,                     "c(-4L)", "LMA-Scaled Tree Size"         , "gdm"   , "log"
                       )#end tribble
   CntTraitCheck = nrow(TraitCheck)

   
   #---~---
   #   Loop through the list of predictors and predictands, and make sure they are all 
   # present in both the look-up table (try_trait) and the training data set (AllomTRY).
   #---~---
   cat0(" + Check if trait/size variable is available.")
   for (u in sequence(CntTraitCheck)){
      # Retrieve check information.
      uName   = TraitCheck$Name [u]
      uIDList = eval(parse(text=TraitCheck$IDList[u]))
      uDesc   = TraitCheck$Desc [u]
      uUnit   = TraitCheck$Unit [u]
      uTrans  = TraitCheck$Trans[u]

      # Figure out if there is any trait that matches this ID.
      IsTrait = which(try_trait$TraitID %in% uIDList)

      # If trait is not in the data base, include it.
      if (length(IsTrait) == 0L){
         cat0("   - Trait ",uName," not available in look-up table. Append it.")

         # Include trait to the look-up table
         try_append = try_template %>%
            mutate( across(matches("Name"   ), ~ uName      )
                  , across(matches("TraitID"), ~ uIDList[1L])
                  , across(matches("Desc"   ), ~ uDesc      )
                  , across(matches("Unit"   ), ~ uUnit      )
                  , across(matches("Trans"  ), ~ uTrans     ) )
         try_trait = rbind(try_trait,try_append)

         # Add dummy vector to AllomTRY
         AllomTRY[[uName]] = NA_real_
      }else{
         uUsed = try_trait$Name[IsTrait]
         cat0("   - Trait ",uName," available (column ",uUsed,"). Ensure it is positively defined.")

         # Make sure the trait is not zero or negative.
         AllomTRY[[uUsed]] = ifelse( test = AllomTRY[[uUsed]] %gt% 0.
                                   , yes  = AllomTRY[[uUsed]]
                                   , no   = NA_real_
                                   )#end ifelse
      }#end if (length(IsTrait) == 0L)
   }#end for (tc in sequence(CntTraitCheck))
   #---~---
   
      
   #---~----
   #   Check if we can fill in wood density. We will try whenever there are some available
   # data, otherwise, we just skip it. In this case, all models using wood density will be
   # automatically rejected.
   #---~---
   # Retrieve column name.
   IsWoodDens = which(try_trait$TraitID %in% c( 4L))
   WoodDens   = try_trait$Name[IsWoodDens][1L]
   if (any(is.finite(AllomTRY[[WoodDens]]))){
      cat0("   - Gap fill wood density.")

      #---~---
      #    Prepare data for filling information for species where no data are available.
      # First we fill in with global sample, then we refine it by sampling data from
      # within the same cluster (when this is possible).
      #---~---
      # 1. Global sampling.
      GlobalIdx       = which(is.finite(AllomTRY[[WoodDens]]))
      FillWoodDens    = sample(x=AllomTRY[[WoodDens]][GlobalIdx],size=nrow(AllomTRY),replace=TRUE)
      # 2. Split reference wood density by cluster.
      ClusterRef      = which(is.finite(AllomTRY[[WoodDens]]) & (! is.na(AllomTRY$Cluster)))
      WDClusterList   = split(x = AllomTRY[[WoodDens]][ClusterRef],f=AllomTRY$Cluster[ClusterRef])
      # 3. Find species that can be filled with cluster data.
      ClusterFill     = ! is.na(AllomTRY$Cluster)
      IdxClusterList  = split(x = ClusterFill, f = AllomTRY$Cluster[ClusterFill])
      # 4. Sample data by cluster
      ClusterWoodDens = 
         mapply( FUN      = function(x,name,ref){
                               ans = sample(x=rep(ref[[name]],times=2L),size=length(x),replace=TRUE)
                               return(ans)
                            }#end function
               , x        = IdxClusterList
               , name     = as.list(names(IdxClusterList))
               , SIMPLIFY = FALSE
               , MoreArgs = list(ref=WDClusterList)
               )#end mapply
      # 5. Update data for filling
      ClusterFill               = c(unlist(IdxClusterList))
      FillWoodDens[ClusterFill] = c(unlist(ClusterWoodDens))
      #    Further improve filling data when species wood density is available
      SpeciesIdx   = match(AllomTRY$ScientificName,SpeciesTRY$ScientificName)
      FillWoodDens = ifelse( test = is.finite(SpeciesTRY[[WoodDens]][SpeciesIdx])
                           , yes  = SpeciesTRY[[WoodDens]][SpeciesIdx]
                           , no   = FillWoodDens
                           )#end ifelse

      #    Fill in data for missing wood density.
      AllomTRY[[WoodDens]] = ifelse( test = is.finite(AllomTRY[[WoodDens]])
                                  , yes  = AllomTRY[[WoodDens]]
                                  , no   = FillWoodDens
                                  )#end ifelse
   }#end if (any(is.finite(AllomTRY[[WoodDens]])))
   #---~---


   #---~----
   #   Check if we can fill in specific leaf area We will try whenever there are some 
   # available data, otherwise, we just skip it. In this case, all models using wood 
   # density will be automatically rejected.
   #---~---
   # Retrieve column name.
   IsSLA = which(try_trait$TraitID %in% c(3086L,3115L,3116L,3117L))[1L]
   SLA   = try_trait$Name[IsSLA]
   if (any(is.finite(AllomTRY[[SLA]]))){
      cat0("   - Gap fill specific leaf area.")
   
      #---~---
      #    Prepare data for filling information for species where no data are available.
      # First we fill in with global sample, then we refine it by sampling data from
      # within the same cluster (when this is possible).
      #---~---
      # 1. Global sampling.
      GlobalIdx       = which(is.finite(AllomTRY[[SLA]]))
      FillSLA         = sample(x=AllomTRY[[SLA]][GlobalIdx],size=nrow(AllomTRY),replace=TRUE)
      # 2. Split reference wood density by cluster.
      ClusterRef      = which(is.finite(AllomTRY[[SLA]]) & (! is.na(AllomTRY$Cluster)))
      WDClusterList   = split(x = AllomTRY[[SLA]][ClusterRef],f=AllomTRY$Cluster[ClusterRef])
      # 3. Find species that can be filled with cluster data.
      ClusterFill     = ! is.na(AllomTRY$Cluster)
      IdxClusterList  = split(x = ClusterFill, f = AllomTRY$Cluster[ClusterFill])
      # 4. Sample data by cluster
      ClusterSLA = 
         mapply( FUN      = function(x,name,ref){
                               ans = sample(x=rep(ref[[name]],times=2L),size=length(x),replace=TRUE)
                               return(ans)
                            }#end function
               , x        = IdxClusterList
               , name     = as.list(names(IdxClusterList))
               , SIMPLIFY = FALSE
               , MoreArgs = list(ref=WDClusterList)
               )#end mapply
      # 5. Update data for filling
      ClusterFill          = c(unlist(IdxClusterList))
      FillSLA[ClusterFill] = c(unlist(ClusterSLA))
      #    Further improve filling data when species wood density is available
      SpeciesIdx = match(AllomTRY$ScientificName,SpeciesTRY$ScientificName)
      FillSLA    = ifelse( test = is.finite(SpeciesTRY[[SLA]][SpeciesIdx])
                         , yes  = SpeciesTRY[[SLA]][SpeciesIdx]
                         , no   = FillSLA
                         )#end ifelse

      #    Fill in data for missing wood density.
      AllomTRY[[SLA]] = ifelse( test = is.finite(AllomTRY[[SLA]])
                              , yes  = AllomTRY[[SLA]]
                              , no   = FillSLA
                              )#end ifelse
   }#end if (any(is.finite(AllomTRY[[SLA]])))

   
   # Find derived quantities
   cat0("   - Calculate derived quantities.")
   DBH      = try_trait$Name[which(try_trait$TraitID %in% c(  21L)                  )[1L]]
   Height   = try_trait$Name[which(try_trait$TraitID %in% c(  18L,3106L)            )[1L]]
   WoodDens = try_trait$Name[which(try_trait$TraitID %in% c(   4L)                  )[1L]]
   SLA      = try_trait$Name[which(try_trait$TraitID %in% c(3086L,3115L,3116L,3117L))[1L]]
   LMA      = try_trait$Name[which(try_trait$TraitID %in% c(  -1L)                  )[1L]]
   Size     = try_trait$Name[which(try_trait$TraitID %in% c(  -2L)                  )[1L]]
   WDSize   = try_trait$Name[which(try_trait$TraitID %in% c(  -3L)                  )[1L]]
   LMSize   = try_trait$Name[which(try_trait$TraitID %in% c(  -4L)                  )[1L]]

   # Leaf mass per area as the inverse of SLA
   AllomTRY[[LMA]] = ifelse( test = AllomTRY[[SLA]] %gt% 0.
                           , yes  = 1./AllomTRY[[SLA]]
                           , no   = NA_real_
                           )#end ifelse

   # Tree Size and the wood-density and SLA scaled versions
   AllomTRY[[Size  ]] = AllomTRY[[DBH     ]]^2 * AllomTRY[[Height]]
   AllomTRY[[WDSize]] = AllomTRY[[WoodDens]]   * AllomTRY[[Size  ]]
   AllomTRY[[LMSize]] = AllomTRY[[LMA     ]]   * AllomTRY[[Size  ]]

   # Save data to an R object.
   cat0(" + Save input data for allometric models to ",basename(rdata_input_allom),".")
   dummy = save( list              = c( "AllomTRY", "try_trait")
               , file              = rdata_input_allom
               , compress          = "xz"
               , compression_level = 9
               )#end save
}#end if (reload_input_allom && file.exists(rdata_input_allom))
```


# Allometric model fitting.

Here we fit a series of allometric models that account for heteroscedasticity when needed. We currently use pre-determined functions for specific traits, but this may be later revisited for more generalised modelling. In all cases, we seek to fit allometric models for each cluster, and a global model, and save the results for deciding whether or not a global model is more parsimonious than the PFT-specific models.  To do this, we use a generic set of functions that will fit multiple allometric models and select the most parsimonious based on the Bayes Information criterion.

## Height allometry

For DBH-Height allometry, we try to fit three models: (1) a modified Michaelis-Mentel relationship, following [Martínez-Cano _et al._, 2020](https://dx.doi.org/10.1111/gcb.15188), (2) a Weibull-based relationship, following [Feldpausch _et_al._, 2012](https://dx.doi.org/10.5194/bg-9-3381-2012), and (3) a simple exponential relationship. We then select the one that has the best predictive power.

```{r, label='try-fitheight',message=FALSE,results='hide',warning=FALSE}
if (reload_height_allom && file.exists(rdata_height_allom)){
   # Reload data
   cat0(" + Reload height allometry.")
   dummy = load(rdata_height_allom)
}else{
   # Set random seed
   if (is.na(rseed_height)){
      SeedPrep    = Sys.time()
      rseed_height = 3600*hour(SeedPrep) + 60*minute(SeedPred) + floor(second(SeedPrep))
   }#end if (is.na(rseed_height))
   dummy = set.seed(rseed_height)

   # Load some files which will likely be updated as the code is developed.
   source(file.path(util_path,"optim.lsq.htscd.r"    ),chdir=TRUE)
   source(file.path(util_path,"TRY_Allometry_Utils.r"),chdir=TRUE)
   
   # Select variables that will be used for deriving the allometry.
   DBH      = try_trait$Name[which(try_trait$TraitID %in% c(21L)      )[1L]]
   Height   = try_trait$Name[which(try_trait$TraitID %in% c(18L,3106L))[1L]]

   # Build tibble with all models to be tested.
   ModelHeight = tribble( ~Model         , ~xName , ~wName        , ~yName
                        , "MartinezCano" , DBH    , NA_character_ , Height
                        , "Weibull"      , DBH    , NA_character_ , Height
                        , "OneLogLinear" , DBH    , NA_character_ , Height
                        )#end tribble
   
   # Find the suite of allometric models and uncertainties
   cat0(" + Fit models relating DBH and height.")
   FitHeight = Allom_Fit( DataTRY         = AllomTRY
                        , try_trait       = try_trait
                        , ModelTRY        = ModelHeight
                        , CategAllom      = CategAllom
                        , UseFixedModel   = UseFixedModel
                        , InfoCrit        = AllomFitXIC
                        , CntAllomMin     = CntAllomMin
                        , CntAllomMax     = CntAllomMax
                        , AllomConfInt    = AllomConfInt
                        , AllomMaxItOptim = AllomMaxItOptim
                        , AllomMaxItGain  = AllomMaxItGain
                        , AllomTolOptim   = AllomTolOptim
                        , AllomTolGain    = AllomTolGain
                        , AllomCntPred    = AllomCntPred
                        , AllomQuantPred  = AllomQuantPred
                        , AllomCntBoot    = AllomCntBoot
                        , UseSizeBins     = UseSizeBins["height"]
                        , xSample         = DBH
                        , xLogSmp         = FALSE
                        , MinSmpPerBin    = MinSmpPerBin
                        , MaxCntBins      = MaxCntBins
                        , Verbose         = TRUE
                        )#end Allom_Fit

   # Save allometric models
   cat0(" + Save fitted allometric models to ",basename(rdata_height_allom),".")
   dummy = save( list              = c( "AllomTRY", "try_trait", "FitHeight")
               , file              = rdata_height_allom
               , compress          = "xz"
               , compression_level = 9
               )#end save
}#end if (reload_height_allom && file.exists(rdata_height_allom))
```




## Aboveground Biomass allometry

For Size-Aboveground Biomass allometry, we try a variation of models based on [Chave _et al._, (2014)](https://dx.doi.org/10.1111/gcb.12629) and [Baker _et al._ (2004)](https://dx.doi.org/10.1111/j.1365-2486.2004.00751.x). Specifically we test whether diameter at breast height ($\mathrm{DBH}$) is sufficient for predicting biomass, or if we should consider a model that uses the volume equivalent ($\mathrm{DBH}^2\,h$), and whether or not to scale the model with wood density. We also try to relate AGB directly to height instead of $\mathrm{DBH}$, which may be easier for lidar applications.


```{r, label='try-fitagb',message=FALSE,results='hide',warning=FALSE}
if (reload_agb_allom && file.exists(rdata_agb_allom)){
   # Reload data
   cat0(" + Reload AGB allometry.")
   dummy = load(rdata_agb_allom)
}else{
   # Set random seed
   if (is.na(rseed_agb)){
      SeedPrep    = Sys.time()
      rseed_height = 3600*hour(SeedPrep) + 60*minute(SeedPred) + floor(second(SeedPrep))
   }#end if (is.na(rseed_agb))
   dummy = set.seed(rseed_agb)

   # Load some files which will likely be updated as the code is developed.
   source(file.path(util_path,"optim.lsq.htscd.r"    ),chdir=TRUE)
   source(file.path(util_path,"TRY_Allometry_Utils.r"),chdir=TRUE)
   
   # Select variables that will be used for deriving the allometry.
   WoodDens = try_trait$Name[which(try_trait$TraitID %in% c(   4L)      )[1L]]
   DBH      = try_trait$Name[which(try_trait$TraitID %in% c(  21L)      )[1L]]
   Height   = try_trait$Name[which(try_trait$TraitID %in% c(  18L,3106L))[1L]]
   Size     = try_trait$Name[which(try_trait$TraitID %in% c(  -2L)      )[1L]]
   WDSize   = try_trait$Name[which(try_trait$TraitID %in% c(  -3L)      )[1L]]
   AGB      = try_trait$Name[which(try_trait$TraitID %in% c(3446L)      )[1L]]

   # Build tibble with all models to be tested.
   ModelAGB = tribble( ~Model         , ~xName , ~wName        , ~yName
                     , "OneLinear"    , Size   , NA_character_ , AGB
                     , "OneLinear"    , WDSize , NA_character_ , AGB
                     , "OneLogLinear" , DBH    , NA_character_ , AGB
                     , "OneLogLinear" , Height , NA_character_ , AGB
                     , "OneLogLinear" , Size   , NA_character_ , AGB
                     , "OneLogLinear" , WDSize , NA_character_ , AGB
                     , "TwoMixLinear" , DBH    , WoodDens      , AGB
                     , "TwoLogLinear" , DBH    , WoodDens      , AGB
                     , "TwoMixLinear" , Height , WoodDens      , AGB
                     , "TwoLogLinear" , Height , WoodDens      , AGB
                     , "TwoMixLinear" , Size   , WoodDens      , AGB
                     , "TwoLogLinear" , Size   , WoodDens      , AGB
                     )#end tribble

   # Find the suite of allometric models and uncertainties
   cat0(" + Fit models relating tree size, specific mass, and AGB.")
   FitAGB = Allom_Fit( DataTRY         = AllomTRY
                     , try_trait       = try_trait
                     , ModelTRY        = ModelAGB
                     , CategAllom      = CategAllom
                     , UseFixedModel   = UseFixedModel
                     , InfoCrit        = AllomFitXIC
                     , CntAllomMin     = CntAllomMin
                     , CntAllomMax     = CntAllomMax
                     , AllomConfInt    = AllomConfInt
                     , AllomMaxItOptim = AllomMaxItOptim
                     , AllomMaxItGain  = AllomMaxItGain
                     , AllomTolOptim   = AllomTolOptim
                     , AllomTolGain    = AllomTolGain
                     , AllomCntPred    = AllomCntPred
                     , AllomQuantPred  = AllomQuantPred
                     , AllomCntBoot    = AllomCntBoot
                     , UseSizeBins     = UseSizeBins["agb"]
                     , xSample         = DBH
                     , xLogSmp         = FALSE
                     , MinSmpPerBin    = MinSmpPerBin
                     , MaxCntBins      = MaxCntBins
                     , Verbose         = TRUE
                     )#end Allom_Fit

   # Save allometric models
   cat0(" + Save fitted allometric models to ",basename(rdata_agb_allom),".")
   dummy = save( list              = c( "AllomTRY", "try_trait", "FitAGB")
               , file              = rdata_agb_allom
               , compress          = "xz"
               , compression_level = 9
               )#end save
}#end if (reload_agb_allom && file.exists(rdata_agb_allom))
```




## Leaf Biomass allometry

For Size-Leaf Biomass allometry, we try to fit models based on [Lescure _et al._ (1983)](http://www.documentation.ird.fr/hor/fdi:010005089),  [Saldarriaga _et al._ (1988)](https://dx.doi.org/10.2307/2260625). Unlike [Saldarriaga _et al._ (1988)](https://dx.doi.org/10.2307/2260625), we do not try to fit a model with different power parameters for each of $\mathrm{DBH}$, height and wood density; instead, we use "tree size" instead. We also fit leaf mass per unit area ($\mathrm{LMA}$) instead of wood density, as LMA is more directly related to leaf biomass and there is evidence of decoupling between wood and leaf traits in tropical forests ([Baraloto et al. (2010)](https://dx.doi.org/10.1111/j.1461-0248.2010.01517.x)). We also try to relate leaf biomass directly to height instead of $\mathrm{DBH}$, which may be easier for lidar applications.

```{r, label='try-fitbleaf',message=FALSE,results='hide',warning=FALSE}
if (reload_bleaf_allom && file.exists(rdata_bleaf_allom)){
   # Reload data
   cat0(" + Reload Leaf biomass allometry.")
   dummy = load(rdata_bleaf_allom)
}else{
   # Set random seed
   if (is.na(rseed_bleaf)){
      SeedPrep    = Sys.time()
      rseed_height = 3600*hour(SeedPrep) + 60*minute(SeedPred) + floor(second(SeedPrep))
   }#end if (is.na(rseed_bleaf))
   dummy = set.seed(rseed_bleaf)

   # Load some files which will likely be updated as the code is developed.
   source(file.path(util_path,"optim.lsq.htscd.r"    ),chdir=TRUE)
   source(file.path(util_path,"TRY_Allometry_Utils.r"),chdir=TRUE)
   
   # Select variables that will be used for deriving the allometry.
   LMA      = try_trait$Name[which(try_trait$TraitID %in% c(  -1L)                  )[1L]]
   SLA      = try_trait$Name[which(try_trait$TraitID %in% c(3086L,3115L,3116L,3117L))[1L]]
   DBH      = try_trait$Name[which(try_trait$TraitID %in% c(  21L)                  )[1L]]
   Height   = try_trait$Name[which(try_trait$TraitID %in% c(  18L,3106L)            )[1L]]
   Size     = try_trait$Name[which(try_trait$TraitID %in% c(  -2L)                  )[1L]]
   LMSize   = try_trait$Name[which(try_trait$TraitID %in% c(  -4L)                  )[1L]]
   BLeaf    = try_trait$Name[which(try_trait$TraitID %in% c( 129L,3441L)            )[1L]]

   # Build tibble with all models to be tested.
   ModelBLeaf = tribble( ~Model         , ~xName , ~wName        , ~yName
                       , "OneLinear"    , Size   , NA_character_ , BLeaf
                       , "OneLinear"    , LMSize , NA_character_ , BLeaf
                       , "OneLogLinear" , DBH    , NA_character_ , BLeaf
                       , "OneLogLinear" , Height , NA_character_ , BLeaf
                       , "OneLogLinear" , Size   , NA_character_ , BLeaf
                       , "OneLogLinear" , LMSize , NA_character_ , BLeaf
#                       , "TwoMixLinear" , DBH    , LMA           , BLeaf
#                       , "TwoLogLinear" , DBH    , SLA           , BLeaf
#                       , "TwoMixLinear" , Height , LMA           , BLeaf
#                       , "TwoLogLinear" , Height , SLA           , BLeaf
#                       , "TwoMixLinear" , Size   , LMA           , BLeaf
#                       , "TwoLogLinear" , Size   , SLA           , BLeaf
                       )#end tribble

   # Find the suite of allometric models and uncertainties
   cat0(" + Fit models relating tree size, specific mass, and leaf biomass.")
   FitBLeaf = Allom_Fit( DataTRY         = AllomTRY
                       , try_trait       = try_trait
                       , ModelTRY        = ModelBLeaf
                       , CategAllom      = CategAllom
                       , UseFixedModel   = UseFixedModel
                       , InfoCrit        = AllomFitXIC
                       , CntAllomMin     = CntAllomMin
                       , CntAllomMax     = CntAllomMax
                       , AllomConfInt    = AllomConfInt
                       , AllomMaxItOptim = AllomMaxItOptim
                       , AllomMaxItGain  = AllomMaxItGain
                       , AllomTolOptim   = AllomTolOptim
                       , AllomTolGain    = AllomTolGain
                       , AllomCntPred    = AllomCntPred
                       , AllomQuantPred  = AllomQuantPred
                       , AllomCntBoot    = AllomCntBoot
                       , UseSizeBins     = UseSizeBins["bleaf"]
                       , xSample         = DBH
                       , xLogSmp         = FALSE
                       , MinSmpPerBin    = MinSmpPerBin
                       , MaxCntBins      = MaxCntBins
                       , Verbose         = TRUE
                       )#end Allom_Fit

   # Save allometric models
   cat0(" + Save fitted allometric models to ",basename(rdata_bleaf_allom),".")
   dummy = save( list              = c( "AllomTRY", "try_trait", "FitBLeaf")
               , file              = rdata_bleaf_allom
               , compress          = "xz"
               , compression_level = 9
               )#end save
}#end if (reload_bleaf_allom && file.exists(rdata_bleaf_allom))
```






## Leaf Area allometry


For Size-Leaf Area allometry, we use similar models as [Lescure _et al._ (1983)](http://www.documentation.ird.fr/hor/fdi:010005089),  [Saldarriaga _et al._ (1988)](https://dx.doi.org/10.2307/2260625), but without trying to scale predictors with wood density or leaf area (akin to  [Longo _et al._, 2020](https://dx.doi.org/10.1029/2020JG005677)). We also try to relate leaf area directly to height instead of $\mathrm{DBH}$, which may be easier for lidar applications.

```{r, label='try-fit-leafarea',message=FALSE,results='hide',warning=FALSE}
if (reload_leaf_area_allom && file.exists(rdata_leaf_area_allom)){
   # Reload data
   cat0(" + Reload LeafArea allometry.")
   dummy = load(rdata_leaf_area_allom)
}else{
   # Set random seed
   if (is.na(rseed_leaf_area)){
      SeedPrep    = Sys.time()
      rseed_height = 3600*hour(SeedPrep) + 60*minute(SeedPred) + floor(second(SeedPrep))
   }#end if (is.na(rseed_leaf_area))
   dummy = set.seed(rseed_leaf_area)

   # Load some files which will likely be updated as the code is developed.
   source(file.path(util_path,"optim.lsq.htscd.r"    ),chdir=TRUE)
   source(file.path(util_path,"TRY_Allometry_Utils.r"),chdir=TRUE)
   
   # Select variables that will be used for deriving the allometry.
   SLA      = try_trait$Name[which(try_trait$TraitID %in% c(3086L,3115L,3116L,3117L))[1L]]
   DBH      = try_trait$Name[which(try_trait$TraitID %in% c( 21L)      )[1L]]
   Height   = try_trait$Name[which(try_trait$TraitID %in% c( 18L,3106L))[1L]]
   Size     = try_trait$Name[which(try_trait$TraitID %in% c( -2L)      )[1L]]
   LeafArea = try_trait$Name[which(try_trait$TraitID %in% c(410L)      )[1L]]

   # Build tibble with all models to be tested. Forcing fit to be with size because this is
   # the best fit for crown area and leaf biomass, and using consisten predictors would 
   # simplify things for FATES.
   ModelLeafArea = tribble( ~Model         , ~xName , ~wName        , ~yName
                          , "OneLinear"    , Size   , NA_character_ , LeafArea
                          , "OneLogLinear" , DBH    , NA_character_ , LeafArea
                          , "OneLogLinear" , Height , NA_character_ , LeafArea
                          , "OneLogLinear" , Size   , NA_character_ , LeafArea
                          )#end tribble

   # Find the suite of allometric models and uncertainties
   cat0(" + Fit models relating tree size and leaf area.")
   FitLeafArea = Allom_Fit( DataTRY         = AllomTRY
                          , try_trait       = try_trait
                          , ModelTRY        = ModelLeafArea
                          , CategAllom      = CategAllom
                          , UseFixedModel   = UseFixedModel
                          , InfoCrit        = AllomFitXIC
                          , CntAllomMin     = CntAllomMin
                          , CntAllomMax     = CntAllomMax
                          , AllomConfInt    = AllomConfInt
                          , AllomMaxItOptim = AllomMaxItOptim
                          , AllomMaxItGain  = AllomMaxItGain
                          , AllomTolOptim   = AllomTolOptim
                          , AllomTolGain    = AllomTolGain
                          , AllomCntPred    = AllomCntPred
                          , AllomQuantPred  = AllomQuantPred
                          , AllomCntBoot    = AllomCntBoot
                          , UseSizeBins     = UseSizeBins["leaf_area"]
                          , xSample         = DBH
                          , xLogSmp         = FALSE
                          , MinSmpPerBin    = MinSmpPerBin
                          , MaxCntBins      = MaxCntBins
                          , Verbose         = TRUE
                          )#end Allom_Fit

   # Save allometric models
   cat0(" + Save fitted allometric models to ",basename(rdata_leaf_area_allom),".")
   dummy = save( list              = c( "AllomTRY", "try_trait", "FitLeafArea")
               , file              = rdata_leaf_area_allom
               , compress          = "xz"
               , compression_level = 9
               )#end save
}#end if (reload_leaf_area_allom && file.exists(rdata_leaf_area_allom))
```




## Crown Area allometry

For Size-Crown Area allometry, we try to fit the exact same models we tried for leaf area, i.e. similar models as [Lescure _et al._ (1983)](http://www.documentation.ird.fr/hor/fdi:010005089),  [Saldarriaga _et al._ (1988)](https://dx.doi.org/10.2307/2260625), but without trying to scale predictors with wood density or leaf area (akin to  [Longo _et al._, 2020](https://dx.doi.org/10.1029/2020JG005677)). We also try to relate crown area directly to height instead of $\mathrm{DBH}$, which may be easier for lidar applications.

```{r, label='try-fit-crownarea',message=FALSE,results='hide',warning=FALSE}
if (reload_crown_area_allom && file.exists(rdata_crown_area_allom)){
   # Reload data
   cat0(" + Reload crown area allometry.")
   dummy = load(rdata_crown_area_allom)
}else{
   # Set random seed
   if (is.na(rseed_crown_area)){
      SeedPrep    = Sys.time()
      rseed_height = 3600*hour(SeedPrep) + 60*minute(SeedPred) + floor(second(SeedPrep))
   }#end if (is.na(rseed_crown_area))
   dummy = set.seed(rseed_crown_area)

   # Load some files which will likely be updated as the code is developed.
   source(file.path(util_path,"optim.lsq.htscd.r"    ),chdir=TRUE)
   source(file.path(util_path,"TRY_Allometry_Utils.r"),chdir=TRUE)

   # Select variables that will be used for deriving the allometry.
   DBH       = try_trait$Name[which(try_trait$TraitID %in% c(21L)      )[1L]]
   Height    = try_trait$Name[which(try_trait$TraitID %in% c(18L,3106L))[1L]]
   Size      = try_trait$Name[which(try_trait$TraitID %in% c(-2L)      )[1L]]
   LMSize    = try_trait$Name[which(try_trait$TraitID %in% c(-4L)      )[1L]]
   WDSize    = try_trait$Name[which(try_trait$TraitID %in% c(-3L)      )[1L]]
   CrownArea = try_trait$Name[which(try_trait$TraitID %in% c(20L)      )[1L]]

   # Build tibble with all models to be tested.
   ModelCrownArea = tribble( ~Model         , ~xName , ~wName        , ~yName
                           , "OneLinear"    , Size   , NA_character_ , CrownArea
                           , "OneLogLinear" , DBH    , NA_character_ , CrownArea
                           , "OneLogLinear" , Height , NA_character_ , CrownArea
                           , "OneLogLinear" , Size   , NA_character_ , CrownArea
                           )#end tribble

   # Find the suite of allometric models and uncertainties
   cat0(" + Fit models relating tree size and crown area.")
   FitCrownArea = Allom_Fit( DataTRY         = AllomTRY
                           , try_trait       = try_trait
                           , ModelTRY        = ModelCrownArea
                           , CategAllom      = CategAllom
                           , UseFixedModel   = UseFixedModel
                           , InfoCrit        = AllomFitXIC
                           , CntAllomMin     = CntAllomMin
                           , CntAllomMax     = CntAllomMax
                           , AllomConfInt    = AllomConfInt
                           , AllomMaxItOptim = AllomMaxItOptim
                           , AllomMaxItGain  = AllomMaxItGain
                           , AllomTolOptim   = AllomTolOptim
                           , AllomTolGain    = AllomTolGain
                           , AllomCntPred    = AllomCntPred
                           , AllomQuantPred  = AllomQuantPred
                           , AllomCntBoot    = AllomCntBoot
                           , UseSizeBins     = UseSizeBins["crown_area"]
                           , xSample         = DBH
                           , xLogSmp         = FALSE
                           , MinSmpPerBin    = MinSmpPerBin
                           , MaxCntBins      = MaxCntBins
                           , Verbose         = TRUE
                           )#end Allom_Fit


   # Save allometric models
   cat0(" + Save fitted allometric models to ",basename(rdata_crown_area_allom),".")
   dummy = save( list              = c( "AllomTRY", "try_trait", "FitCrownArea")
               , file              = rdata_crown_area_allom
               , compress          = "xz"
               , compression_level = 9
               )#end save
}#end if (reload_crown_area_allom && file.exists(rdata_crown_area_allom))
```





## Crown Depth allometry

For Height-Crown Depth allometry, we fit a model similar to [Poorter et al. (2006)](https://dx.doi.org/10.1890/0012-9658(2006)87{$[$}1289:AOMTST{$]$}2.0.CO;2), using height as predictor. We also try to fit a purely linear function.

```{r, label='try-fit-crowndepth',message=FALSE,results='hide',warning=FALSE}
if (reload_crown_depth_allom && file.exists(rdata_crown_depth_allom)){
   # Reload data
   cat0(" + Reload crown depth allometry.")
   dummy = load(rdata_crown_depth_allom)
}else{
   # Set random seed
   if (is.na(rseed_crown_depth)){
      SeedPrep    = Sys.time()
      rseed_height = 3600*hour(SeedPrep) + 60*minute(SeedPred) + floor(second(SeedPrep))
   }#end if (is.na(rseed_crown_depth))
   dummy = set.seed(rseed_crown_depth)

   # Load some files which will likely be updated as the code is developed.
   source(file.path(util_path,"optim.lsq.htscd.r"    ),chdir=TRUE)
   source(file.path(util_path,"TRY_Allometry_Utils.r"),chdir=TRUE)
   
   # Select variables that will be used for deriving the allometry.
   Height     = try_trait$Name[which(try_trait$TraitID %in% c( 18L,3106L))[1L]]
   CrownDepth = try_trait$Name[which(try_trait$TraitID %in% c(773L)      )[1L]]

   # Build tibble with all models to be tested.
   ModelCrownDepth = tribble( ~Model         , ~xName , ~wName        , ~yName
                            , "OneLinear"    , Height , NA_character_ , CrownDepth
                            , "OneLogLinear" , Height , NA_character_ , CrownDepth
                            )#end tribble

   # Find the suite of allometric models and uncertainties
   cat0(" + Fit models relating tree height and crown depth.")
   FitCrownDepth = Allom_Fit( DataTRY         = AllomTRY
                            , try_trait       = try_trait
                            , ModelTRY        = ModelCrownDepth
                            , CategAllom      = CategAllom
                            , UseFixedModel   = UseFixedModel
                            , InfoCrit        = AllomFitXIC
                            , CntAllomMin     = CntAllomMin
                            , CntAllomMax     = CntAllomMax
                            , AllomConfInt    = AllomConfInt
                            , AllomMaxItOptim = AllomMaxItOptim
                            , AllomMaxItGain  = AllomMaxItGain
                            , AllomTolOptim   = AllomTolOptim
                            , AllomTolGain    = AllomTolGain
                            , AllomCntPred    = AllomCntPred
                            , AllomQuantPred  = AllomQuantPred
                            , AllomCntBoot    = AllomCntBoot
                            , UseSizeBins     = UseSizeBins["crown_depth"]
                            , xSample         = Height
                            , xLogSmp         = FALSE
                            , MinSmpPerBin    = MinSmpPerBin
                            , MaxCntBins      = MaxCntBins
                            , Verbose         = TRUE
                            )#end Allom_Fit

   # Save allometric models
   cat0(" + Save fitted allometric models to ",basename(rdata_crown_depth_allom),".")
   dummy = save( list              = c( "AllomTRY", "try_trait", "FitCrownDepth")
               , file              = rdata_crown_depth_allom
               , compress          = "xz"
               , compression_level = 9
               )#end save
}#end if (reload_crown_depth_allom && file.exists(rdata_crown_depth_allom))
```


## Bark Thickness allometry

For bark thickness allometry, we seek for relationships with either DBH or height, potentially scaled by wood density.

```{r, label='try-fit-barkthick',message=FALSE,results='hide',warning=FALSE}
if (reload_bark_thick_allom && file.exists(rdata_bark_thick_allom)){
   # Reload data
   cat0(" + Reload bark thickness allometry.")
   dummy = load(rdata_bark_thick_allom)
}else{
   # Set random seed
   if (is.na(rseed_bark_thick)){
      SeedPrep    = Sys.time()
      rseed_height = 3600*hour(SeedPrep) + 60*minute(SeedPred) + floor(second(SeedPrep))
   }#end if (is.na(rseed_bark_thick))
   dummy = set.seed(rseed_bark_thick)

   # Load some files which will likely be updated as the code is developed.
   source(file.path(util_path,"optim.lsq.htscd.r"    ),chdir=TRUE)
   source(file.path(util_path,"TRY_Allometry_Utils.r"),chdir=TRUE)
   
   # Select variables that will be used for deriving the allometry.
   DBH       = try_trait$Name[which(try_trait$TraitID %in% c(  21L)      )[1L]]
   Height    = try_trait$Name[which(try_trait$TraitID %in% c(  18L,3106L))[1L]]
   Size      = try_trait$Name[which(try_trait$TraitID %in% c(  -2L)      )[1L]]
   BarkThick = try_trait$Name[which(try_trait$TraitID %in% c(  24L)      )[1L]]

   # Build tibble with all models to be tested.
   ModelBarkThick = tribble( ~Model         , ~xName , ~wName        , ~yName
                           , "OneLinear"    , DBH    , NA_character_ , BarkThick
                           , "OneLinear"    , Height , NA_character_ , BarkThick
                           , "OneLinear"    , Size   , NA_character_ , BarkThick
                           , "OneLogLinear" , DBH    , NA_character_ , BarkThick
                           , "OneLogLinear" , Height , NA_character_ , BarkThick
                           , "OneLogLinear" , Size   , NA_character_ , BarkThick
                           )#end tribble

   # Find the suite of allometric models and uncertainties
   cat0(" + Fit models relating tree size and bark thickness.")
   FitBarkThick = Allom_Fit( DataTRY         = AllomTRY
                           , try_trait       = try_trait
                           , ModelTRY        = ModelBarkThick
                           , CategAllom      = CategAllom
                           , UseFixedModel   = UseFixedModel
                           , InfoCrit        = AllomFitXIC
                           , CntAllomMin     = CntAllomMin
                           , CntAllomMax     = CntAllomMax
                           , AllomConfInt    = AllomConfInt
                           , AllomMaxItOptim = AllomMaxItOptim
                           , AllomMaxItGain  = AllomMaxItGain
                           , AllomTolOptim   = AllomTolOptim
                           , AllomTolGain    = AllomTolGain
                           , AllomCntPred    = AllomCntPred
                           , AllomQuantPred  = AllomQuantPred
                           , AllomCntBoot    = AllomCntBoot
                           , UseSizeBins     = UseSizeBins["bark_thick"]
                           , xSample         = DBH
                           , xLogSmp         = FALSE
                           , MinSmpPerBin    = MinSmpPerBin
                           , MaxCntBins      = MaxCntBins
                           , Verbose         = TRUE
                           )#end Allom_Fit

   # Save allometric models
   cat0(" + Save fitted allometric models to ",basename(rdata_bark_thick_allom),".")
   dummy = save( list              = c( "AllomTRY", "try_trait", "FitBarkThick")
               , file              = rdata_bark_thick_allom
               , compress          = "xz"
               , compression_level = 9
               )#end save
}#end if (reload_bark_thick_allom && file.exists(rdata_bark_thick_allom))
```



# Prepare data for plotting

Here we define the list of predicted variables along with the description, units and the associated object with the fitted model.
```{r, label='plot-schedule',message=FALSE,results='hide',warning=FALSE}

# Select predictands of the various allometry models
idHeight     = try_trait$TraitID[which(try_trait$TraitID %in% c(  18L,3106L))[1L]]
idAGB        = try_trait$TraitID[which(try_trait$TraitID %in% c(3446L)      )[1L]]
idBLeaf      = try_trait$TraitID[which(try_trait$TraitID %in% c( 129L,3441L))[1L]]
idLeafArea   = try_trait$TraitID[which(try_trait$TraitID %in% c( 410L)      )[1L]]
idCrownArea  = try_trait$TraitID[which(try_trait$TraitID %in% c(  20L)      )[1L]]
idCrownDepth = try_trait$TraitID[which(try_trait$TraitID %in% c( 773L)      )[1L]]
idBarkThick  = try_trait$TraitID[which(try_trait$TraitID %in% c(  24L)      )[1L]]

# List of objects to fit models
PlotTRY = tribble( ~TraitID    , ~DefName
                 , idHeight    , "Height"
                 , idAGB       , "AGB"
                 , idBLeaf     , "BLeaf"
                 , idLeafArea  , "LeafArea"
                 , idCrownArea , "CrownArea"
                 , idCrownDepth, "CrownDepth"
                 , idBarkThick , "BarkThick"
                 )#end tribble


# Keep only plots that are associated with existing objects
PlotTRY = PlotTRY %>%
   mutate( FitName = paste0("Fit",DefName)) %>%
   filter(FitName %in% ls(envir=.GlobalEnv))

# Number of models to plot
CntPlotTRY = nrow(PlotTRY)
```




# Plot the allometric relationships.

## Goodness-of-fit assessment

First, we plot the model goodness of fit for every variable of interest, by plotting scatter plots of predicted estimates against observed.

```{r,label='plot-scatter-predobs',message=FALSE,results='hide',warning=FALSE,fig.width=8, fig.height=6}

# Load some files which will likely be updated as the code is developed.
source(file.path(util_path,"TRY_Allometry_Utils.r"),chdir=TRUE)
source(file.path(util_path,"unitlist.r"           ),chdir=TRUE)



# Find colours and levels
if (DensPalette %in% BrewerPalInfo){
   DensColours = RColorBrewer::brewer.pal(n=5,name=DensPalette)
}else if (h_csch %in% viridis_pal_info){
   DensColours = viridis::viridis(n=5,option=DensPalette)
}else{
   DensPalette = match.fun(DensPalette)
   DensColours = DensPalette(n=5)
}#end if (v_cnorm %in% brewer_pal_info)

# Reverse palette in case it is needed
if (DensReverse) DensColours = rev(DensColours)

# Create colour ramp palette
DensRamp = grDevices::colorRampPalette(colors=DensColours,space="Lab")


# Initialise list of output variables
gg_allom = list()

# Loop through all models
GoodLoop = sequence(CntPlotTRY * plot_allom_scatter)
for (p in GoodLoop){
   # Retrieve name containing the model fitting object and the trait ID for predictand.
   y        = match(PlotTRY$TraitID[p],try_trait$TraitID)
   yTraitID = try_trait$TraitID[y]
   yDefName = PlotTRY$DefName  [p]
   yName    = try_trait$Name   [y]
   yDesc    = try_trait$Desc   [y]
   yDescObs = paste0(yDesc," - Observed")
   yDescMod = paste0(yDesc," - Fitted"  )
   yUnit    = try_trait$Unit   [y]
   yTrans   = if(yDefName %in% names(yTransAllom)){yTransAllom[yDefName]}else{"identity"}
   yTPlot   = switch( EXPR = yTrans, log = "log10", neglog="neglog10",yTrans)
   yFit     = get(PlotTRY$FitName[p])
   ySumm    = yFit$SummAllom
   
   cat0(" + Plot scatter plot of predicted allometry for ",yDesc,".")

   
   # Initialise data object that will contain predictors, predicted value and observed
   PredTRY = tibble( Class      = character(0L)
                   , TraitClass = character(0L)
                   , x          = numeric(0L)
                   , w          = numeric(0L)
                   , yObs       = numeric(0L)
                   , yMod       = numeric(0L)
                   )#end tibble

     
   # Find the number of traits for plotting the model
   for (ct in sequence(CntCategAllom)){
      # Load settings for the classes.
      zTrait = CategAllom$TraitID   [ct]
      zClass = CategAllom$Name      [ct]
      zName  = CategAllom$TraitClass[ct]
      zDesc  = CategAllom$DescClass [ct]

      # Select model and parameters
      zIdx   = which( (ySumm$TraitClass %in% zName) & (ySumm$Class %in% zClass) )
      zModel = ySumm$Function[zIdx]
      zParam = c( a0 = ySumm$a0[zIdx]
                , a1 = ySumm$a1[zIdx]
                , a2 = ySumm$a2[zIdx]
                , s0 = ySumm$s0[zIdx]
                , s1 = ySumm$s1[zIdx]
                )#end c
      xName  = ySumm$xName[zIdx]
      wName  = ySumm$wName[zIdx]

      # Select data from the allometry data set
      if (is.na(zTrait)){
         zSel = rep(x=TRUE,times=nrow(AllomTRY))
      }else{
         zSel = AllomTRY[[zName]] %in% zClass
      }#end if (is.na(zTrait))

      # Initialise data from the allometry data set. 
      PredNow = tibble( Class      = zClass
                      , TraitClass = zName
                      , x          = if(is.na(xName)){NA_real_}else{AllomTRY[[xName]][zSel]}
                      , w          = if(is.na(wName)){NA_real_}else{AllomTRY[[wName]][zSel]}
                      , yObs       = if(is.na(yName)){NA_real_}else{AllomTRY[[yName]][zSel]}
                      , yMod       = if(is.na(zModel)){
                                        NA_real_
                                     }else{
                                        Allom_Pred( x     = x
                                                  , w     = w
                                                  , param = zParam
                                                  , fun   = zModel
                                                  , ans   = "mu"
                                                  )#end try TRY_AllomPred
                                     }#end if (is.na(zModel)) 
                      )#end tibble
      # Bind predictions to a unified object
      PredTRY = as_tibble(rbind(PredTRY,PredNow))
   }#end for (ct in seq_along(CategTrait))


   # Find the number of traits for plotting the model
   CategSel    = is.finite(CategAllom$TraitID)
   CategTrait  = sort(unique(CategAllom$TraitID[CategSel]))
   for (ct in seq_along(CategTrait)){
      # Load settings for the classes.
      zTrait = CategTrait[ct]
      if (zTrait %in% 0L){
         zName  = "Cluster"
         zDesc  = "Data-based cluster class"
      }else{
         z      = match(zTrait,try_trait$TraitID)
         zName  = try_trait$Name   [z]
         zDesc  = try_trait$Desc   [z]
      }#end if (zTrait %in% 0L)
      zTitle = paste0("Allometric model for ",yDesc," by ",tolower(zDesc))

            
      # Set path for this group of plots.
      categ_allom_path = file.path(allom_path,zName)
      
      # Select categories for this trait type.
      CategNow = CategAllom %>%
         rename( Class = Name) %>%
         filter( (TraitID %in% zTrait) | is.na(TraitID)) %>%
         filter(! Class %in% "UKN") %>%
         filter(! duplicated(Class))
      CntCategNow = nrow(CategNow)

      
      # Find the optimal number of rows and columns, and find multiplication factor 
      # so plots are not ridiculously small
      nColCateg   = min(3,ceiling(sqrt(CntCategNow)))
      nRowCateg   = ceiling(CntCategNow/nColCateg)
      multHeight  = sqrt(nRowCateg)
      multWidth   = multHeight * nColCateg / nRowCateg


      # Select valid data for the axes
      oSel = switch( EXPR     = yTrans
                   , identity = is.finite(PredTRY$yObs)
                   , log      = PredTRY$yObs %gt% 0.
                   , neglog   = PredTRY$yObs %lt% 0.
                   , sqrt     = PredTRY$yObs %ge% 0.
                   , cbrt     = is.finite(PredTRY$yObs)
                   , stop(paste0(" Invalid transformation for trait ",yName," (TraitID = ",yTraitID,")."))
                   )#end switch
      mSel = switch( EXPR     = yTrans
                   , identity = is.finite(PredTRY$yMod)
                   , log      = PredTRY$yMod %gt% 0.
                   , neglog   = PredTRY$yMod %lt% 0.
                   , sqrt     = PredTRY$yMod %ge% 0.
                   , cbrt     = is.finite(PredTRY$yMod)
                   , stop(paste0(" Invalid transformation for trait ",yName," (TraitID = ",yTraitID,")."))
                   )#end switch

      
      #    Subset data. We keep only valid data points that were either assigned one of the
      # categories of this class, or classified as "All".
      PredPlot    = PredTRY %>%
         filter(oSel & mSel) %>%
         filter( Class %in% CategNow$Class) %>%
         mutate( Class = factor(x=Class,levels=CategNow$Class) )
      CntPredPlot = nrow(PredPlot)
      PlotAllom   = CntPredPlot %ge% CntAllomMin
      PlotFit     = any(is.finite(PredPlot$yMod))

      # Decide whether to use density plots or scatter plots for observations.
      DensObserv = ( CntPredPlot %ge% CntDensThresh ) && ( yTrans %in% "identity" )

      # Plot only the variables with meaningful data
      if (PlotAllom){
         cat0("   - Plot the allometric model for ",yDesc,", by ",zDesc,".")

         
         # Set path for this group of plots.
         dummy = dir.create(categ_allom_path,recursive=TRUE,showWarnings=FALSE)
 

         # Find limits for plots.   
         xyLimit = range(c(PredPlot$yObs,PredPlot$yMod),finite=TRUE)
         xyWidth = diff(xyLimit) / CntDensBin

         # Set categories for colours, lines and symbols
         colClasses         = CategNow$Colour
         names(colClasses)  = CategNow$Class
         pchClasses         = CategNow$Symbol
         names(pchClasses)  = CategNow$Class
         
         
         # Build the plot
         gg_now = ggplot()
         gg_now = gg_now + facet_wrap( ~ Class, ncol = nColCateg, labeller = label_parsed)
         if (DensObserv){
            gg_now = gg_now + geom_bin_2d(data=PredPlot,aes(x=yMod,y=yObs),binwidth=c(xyWidth,xyWidth))
            gg_now = gg_now + scale_fill_gradientn( name    = "Point density"
                                                  , colours = DensRamp(n=DensCntColour)
                                                  , limits  = DensRange
                                                  , labels  = label_number()
                                                  , trans   = DensTrans
                                                  )#end scale_fill_gradientn
         }else{
            gg_now = gg_now + geom_point( data    = PredPlot
                                        , mapping = aes(x=yMod,y=yObs)
                                        , colour  = PointColour
                                        , shape   = PointShape
                                        , size    = PointSize
                                        )#end geom_point
         }#end if (DensObserv)

         if (PlotFit){
            gg_now = gg_now + geom_abline( data        = PredPlot
                                         , mapping     = aes(colour=Class) 
                                         , slope       = 1.0
                                         , intercept   = 0.0
                                         , linewidth   = 1.5
                                         , linetype    = "solid"
                                         , inherit.aes = FALSE
                                         , show.legend = FALSE
                                         )#end geom_line
            gg_now = gg_now + scale_colour_manual(values=colClasses)
            gg_now = gg_now + labs( x        = desc.unit(desc=yDescMod,unit=untab[[yUnit]])
                                  , y        = desc.unit(desc=yDescObs,unit=untab[[yUnit]])
                                  , colour   = element_blank()
                                  , fill     = element_blank()
                                  , shape    = element_blank()
                                  , linetype = element_blank()
                                  , title    = zTitle
                                  , subtitle = LabelSubtitle
                                  )#end labs
         }else{
            gg_now = gg_now + labs( x        = desc.unit(desc=yDescMod,unit=untab[[yUnit]])
                                  , y        = desc.unit(desc=yDescObs,unit=untab[[yUnit]])
                                  , colour   = element_blank()
                                  , shape    = element_blank()
                                  , title    = zTitle
                                  , subtitle = LabelSubtitle
                                  )#end labs
         }#end if (PlotFit)
         gg_now = gg_now + theme_grey( base_size = gg_ptsz, base_family = "Helvetica",base_line_size = 0.8,base_rect_size =0.8)
         gg_now = gg_now + theme( axis.text.x       = element_text( size = gg_ptsz, margin = unit(rep(0.35,times=4),"cm"))
                                , axis.text.y       = element_text( size = gg_ptsz, margin = unit(rep(0.35,times=4),"cm"))
                                , plot.title        = element_text( size = gg_ptsz)
                                , plot.subtitle     = element_text( size = 0.7*gg_ptsz)
                                , axis.ticks.length = unit(-0.25,"cm")
                                , legend.position   = "bottom"
                                , legend.direction  = "horizontal"
                                )#end theme
         gg_now = gg_now + guides(fill= guide_colourbar(barwidth=unit(1./6.,"npc")))
         gg_now = gg_now + scale_x_continuous(trans=yTPlot,limits=xyLimit)
         gg_now = gg_now + scale_y_continuous(trans=yTPlot,limits=xyLimit)

         # Save plot in every format requested.
         for (d in sequence(ndevice)){
            f_output = paste0("AllomScatter_",yName,"_by-",zName,"_",base_suffix,".",gg_device[d])
            dummy    = ggsave( filename = f_output
                             , plot     = gg_now
                             , device   = gg_device[d]
                             , path     = categ_allom_path
                             , width    = gg_square * multWidth
                             , height   = gg_square * multHeight
                             , units    = gg_units
                             , dpi      = gg_depth
                             )#end ggsave
         }#end for (o in sequence(nout))
         
         # Write plot settings to the list.
         yzName             = paste0(yName,"_",zName)
         gg_allom[[yzName]] = gg_now
      }else{
         cat0(" + Skip plot for ",yDesc," by ",zDesc,": too few valid points.")
      }#end if (PlotAllom)
   }#end for (ct in seq_along(CategTrait))
}#end for (p in GoodLoop){

# If sought, plot images on screen
cnt_gg_allom = length(gg_allom)
if (gg_screen && (cnt_gg_allom %gt% 0L)){
   gg_show = sort(sample.int(n=cnt_gg_allom,size=min(cnt_gg_allom,3L),replace=FALSE))
   gg_allom[gg_show]
}#end if (gg_screen && (length(gg_allom) %gt% 0L))
```

## Functional form.

In this block, we plot the functional shape of the fitted equations, and compare the results for the different categories. We only plot this if all classes are using the same functional form.

```{r,label='plot-fitted-function',message=FALSE,results='hide',warning=FALSE,fig.width=8, fig.height=6}

# Load some files which will likely be updated as the code is developed.
source(file.path(util_path,"TRY_Allometry_Utils.r"),chdir=TRUE)
source(file.path(util_path,"unitlist.r"           ),chdir=TRUE)


# Initialise list of output variables
gg_funct = list()

# Loop through all models
FunLoop = sequence(CntPlotTRY * plot_allom_fun * UseFixedModel)
for (p in FunLoop){
   # Retrieve name containing the model fitting object and the trait ID for predictand.
   y        = match(PlotTRY$TraitID[p],try_trait$TraitID)
   yTraitID = try_trait$TraitID[y]
   yDefName = PlotTRY$DefName  [p]
   yName    = try_trait$Name   [y]
   yDesc    = try_trait$Desc   [y]
   yDescObs = paste0(yDesc," - Observed")
   yDescMod = paste0(yDesc," - Fitted"  )
   yUnit    = try_trait$Unit   [y]
   yTrans   = if(yDefName %in% names(yTransAllom)){yTransAllom[yDefName]}else{"identity"}
   yTPlot   = switch( EXPR = yTrans, log = "log10", neglog="neglog10",yTrans)
   yFit     = get(PlotTRY$FitName[p])
   ySumm    = yFit$SummAllom
   
   cat0(" + Plot scatter plot of predicted allometry for ",yDesc,".")

   
   # Copy data object that contains the fitted model.
   PredTRY = yFit$PredAllom

   # Find the number of traits for plotting the model
   CategSel    = is.finite(CategAllom$TraitID)
   CategTrait  = sort(unique(CategAllom$TraitID[CategSel]))
   for (ct in seq_along(CategTrait)){
      # Load settings for the classes.
      zTrait = CategTrait[ct]
      if (zTrait %in% 0L){
         zName  = "Cluster"
         zDesc  = "Data-based cluster class"
      }else{
         z      = match(zTrait,try_trait$TraitID)
         zName  = try_trait$Name   [z]
         zDesc  = try_trait$Desc   [z]
      }#end if (zTrait %in% 0L)
      zTitle = paste0("Allometric model for ",yDesc," by ",tolower(zDesc))

      # Select model and parameters
      zIdx   = which( ( ySumm$TraitClass %in% zName ) & (! is.na(ySumm$Function) ) )[1L]
      zModel = ySumm$Function[zIdx]
      xName  = ySumm$xName   [zIdx]
      wName  = ySumm$wName   [zIdx]

      # Load information for X axis.
      x        = match(xName,try_trait$Name)
      xTraitID = try_trait$TraitID[x]
      xLabel   = sprintf("ID_%+4.4i",xTraitID)
      xLabel   = gsub(pattern="\\+",replacement="p",x=xLabel)
      xLabel   = gsub(pattern="\\-",replacement="m",x=xLabel)
      xDefName = switch( EXPR     = xLabel
                       , ID_m0004 = "LMSize"
                       , ID_m0003 = "WDSize"
                       , ID_m0002 = "Size"
                       , ID_p0018 = "Height"
                       , ID_p0021 = "DBH"
                       , ID_p3106 = "Height"
                       )#end switch
      xName    = try_trait$Name   [x]
      xDesc    = try_trait$Desc   [x]
      xUnit    = try_trait$Unit   [x]
      xTrans   = if(xDefName %in% names(xTransAllom)){xTransAllom[xDefName]}else{"identity"}
      xTPlot   = switch( EXPR = xTrans, log = "log10", neglog="neglog10",xTrans)
      
            
      # Set path for this group of plots.
      categ_allom_path = file.path(allom_path,zName)
      
      # Select categories for this trait type.
      CategNow = CategAllom %>%
         rename( Class = Name) %>%
         filter( (TraitID %in% zTrait) | is.na(TraitID)) %>%
         filter(! Class %in% "UKN") %>%
         filter(! duplicated(Class))
      CntCategNow = nrow(CategNow)

      
      # Find the optimal number of rows and columns, and find multiplication factor 
      # so plots are not ridiculously small
      nColCateg   = min(3,ceiling(sqrt(CntCategNow)))
      nRowCateg   = ceiling(CntCategNow/nColCateg)
      multHeight  = sqrt(nRowCateg)
      multWidth   = multHeight * nColCateg / nRowCateg


      # Select valid data for the axes
      xSel = switch( EXPR     = xTrans
                   , identity = is.finite(PredTRY$x)
                   , log      = PredTRY$x %gt% 0.
                   , neglog   = PredTRY$x %lt% 0.
                   , sqrt     = PredTRY$x %ge% 0.
                   , cbrt     = is.finite(PredTRY$x)
                   , stop(paste0(" Invalid transformation for trait ",xName," (TraitID = ",xTraitID,")."))
                   )#end switch
      ySel = switch( EXPR     = yTrans
                   , identity = is.finite(PredTRY$y)
                   , log      = PredTRY$y %gt% 0.
                   , neglog   = PredTRY$y %lt% 0.
                   , sqrt     = PredTRY$y %ge% 0.
                   , cbrt     = is.finite(PredTRY$y)
                   , stop(paste0(" Invalid transformation for trait ",yName," (TraitID = ",yTraitID,")."))
                   )#end switch

      
      #    Subset data. We keep only valid data points that were either assigned one of the
      # categories of this class, or classified as "All".
      PredPlot    = PredTRY %>%
         filter(xSel & ySel) %>%
         filter( Class %in% CategNow$Class) %>%
         mutate( Class = factor(x=Class,levels=CategNow$Class) )

      #   Make sure all classes are included. In case any class is missing, add a dummy data point with
      # no information.
      TallyClass = table(PredPlot$Class)
      MissClass  = names(TallyClass[TallyClass == 0L])
      if (length(MissClass) > 0L){
           mIdx = match(MissClass,CategNow$Class)

           # Create a single line for missing categories, but leave no data.
           PredMiss = PredPlot[rep(1L,times=length(MissClass)),,drop=FALSE] %>%
              mutate(across(where(is.numeric), ~ .x * NA_real_)) %>%
              mutate( Class      = factor(x=CategNow$Class[mIdx],levels=CategNow$Class)
                    , TraitClass = CategNow$TraitClass[mIdx]
                    , DescClass  = CategNow$DescClass [mIdx]
                    )#end mutate

           # Append dummy data frame
           PredPlot = rbind(PredPlot,PredMiss)
      }#end if (length(MissClass) > 0L)
      
      
      # Decide whether or not to plot the data.
      CntPredPlot = nrow(PredPlot)
      PlotAllom   = CntPredPlot %ge% CntAllomMin
      PlotFit     = any(is.finite(PredPlot$y))

      # Plot only the variables with meaningful data
      if (PlotAllom){
         cat0("   - Plot the allometric model for ",yDesc,", by ",zDesc,".")

         
         # Set path for this group of plots.
         dummy = dir.create(categ_allom_path,recursive=TRUE,showWarnings=FALSE)
 

         # Find limits for plots.   
         xLimit = range(c(PredPlot$x),finite=TRUE)
         yLimit = range(c(PredPlot$y,PredPlot$yLwr,PredPlot$yUpr),finite=TRUE)

         # Set categories for colours, lines and symbols
         colClasses         = CategNow$Colour
         names(colClasses)  = CategNow$Class
         pchClasses         = CategNow$Symbol
         names(pchClasses)  = CategNow$Class
         labClasses         = names(colClasses)
         
         
         # Build the plot
         gg_now = ggplot()
         if (PlotFit){
            gg_now = gg_now + scale_colour_manual(name="",aesthetics="fill"  ,labels=labClasses,values=colClasses)
            gg_now = gg_now + scale_colour_manual(name="",aesthetics="colour",labels=labClasses,values=colClasses)
            gg_now = gg_now + geom_ribbon( data        = PredPlot
                                         , mapping     = aes(x=x,ymin=yLwr,ymax=yUpr,fill=Class)
                                         , alpha       = 0.25
                                         , colour      = "transparent"
                                         , linewidth   = 1.0
                                         , linetype    = "dotdash"
                                         , inherit.aes = FALSE
                                         , show.legend = FALSE
                                         )#end geom_line
            gg_now = gg_now + geom_line( data        = PredPlot
                                       , mapping     = aes(x=x,y=y,colour=Class)
                                       , linewidth   = 1.5
                                       , linetype    = "solid"
                                       , inherit.aes = FALSE
                                       , show.legend = TRUE
                                       )#end geom_line
            gg_now = gg_now + labs( x        = desc.unit(desc=xDesc,unit=untab[[xUnit]])
                                  , y        = desc.unit(desc=yDesc,unit=untab[[yUnit]])
                                  , colour   = element_blank()
                                  , fill     = element_blank()
                                  , shape    = element_blank()
                                  , linetype = element_blank()
                                  , title    = zTitle
                                  , subtitle = LabelSubtitle
                                  )#end labs
         }else{
            gg_now = gg_now + labs( x        = desc.unit(desc=xDesc,unit=untab[[xUnit]])
                                  , y        = desc.unit(desc=yDesc,unit=untab[[yUnit]])
                                  , colour   = element_blank()
                                  , shape    = element_blank()
                                  , title    = zTitle
                                  , subtitle = LabelSubtitle
                                  )#end labs
         }#end if (PlotFit)
         gg_now = gg_now + theme_grey( base_size = gg_ptsz, base_family = "Helvetica",base_line_size = 0.8,base_rect_size =0.8)
         gg_now = gg_now + theme( axis.text.x       = element_text( size = gg_ptsz, margin = unit(rep(0.35,times=4),"cm"))
                                , axis.text.y       = element_text( size = gg_ptsz, margin = unit(rep(0.35,times=4),"cm"))
                                , plot.title        = element_text( size = gg_ptsz)
                                , plot.subtitle     = element_text( size = 0.7*gg_ptsz)
                                , axis.ticks.length = unit(-0.25,"cm")
                                , legend.position   = "bottom"
                                , legend.direction  = "horizontal"
                                )#end theme
         gg_now = gg_now + guides(fill= guide_colourbar(barwidth=unit(1./6.,"npc")))
         gg_now = gg_now + scale_x_continuous(trans=xTPlot,limits=xLimit)
         gg_now = gg_now + scale_y_continuous(trans=yTPlot,limits=yLimit)

         # Save plot in every format requested.
         for (d in sequence(ndevice)){
            f_output = paste0("AllomFit_",yName,"_by-",zName,"_",base_suffix,".",gg_device[d])
            dummy    = ggsave( filename = f_output
                             , plot     = gg_now
                             , device   = gg_device[d]
                             , path     = categ_allom_path
                             , width    = gg_square
                             , height   = gg_square
                             , units    = gg_units
                             , dpi      = gg_depth
                             )#end ggsave
         }#end for (o in sequence(nout))
         
         # Write plot settings to the list.
         yzName             = paste0(yName,"_",zName)
         gg_funct[[yzName]] = gg_now
      }else{
         cat0(" + Skip plot for ",yDesc," by ",zDesc,": too few valid points.")
      }#end if (PlotAllom)
   }#end for (ct in seq_along(CategTrait))
}#end for (p in GoodLoop){

# If sought, plot images on screen
cnt_gg_funct = length(gg_funct)
if (gg_screen && (cnt_gg_funct %gt% 0L)){
   gg_show = sort(sample.int(n=cnt_gg_funct,size=min(cnt_gg_funct,3L),replace=FALSE))
   gg_funct[gg_show]
}#end if (gg_screen && (cnt_gg_funct %gt% 0L))
```



## Functional form with actual points.

In this block, we plot the functional shape of the fitted equations, and compare the results for the different categories. We only plot this if all classes are using the same functional form.

```{r,label='plot-fitted-scatter',message=FALSE,results='hide',warning=FALSE,fig.width=8, fig.height=6}

# Load some files which will likely be updated as the code is developed.
source(file.path(util_path,"TRY_Allometry_Utils.r"),chdir=TRUE)
source(file.path(util_path,"unitlist.r"           ),chdir=TRUE)


# Initialise list of output variables
gg_funct = list()

# Loop through all models
FunLoop = sequence(CntPlotTRY * plot_allom_fun * UseFixedModel)
for (p in FunLoop){
   # Retrieve name containing the model fitting object and the trait ID for predictand.
   y        = match(PlotTRY$TraitID[p],try_trait$TraitID)
   yTraitID = try_trait$TraitID[y]
   yDefName = PlotTRY$DefName  [p]
   yName    = try_trait$Name   [y]
   yDesc    = try_trait$Desc   [y]
   yDescObs = paste0(yDesc," - Observed")
   yDescMod = paste0(yDesc," - Fitted"  )
   yUnit    = try_trait$Unit   [y]
   yTrans   = if(yDefName %in% names(yTransAllom)){yTransAllom[yDefName]}else{"identity"}
   yTPlot   = switch( EXPR = yTrans, log = "log10", neglog="neglog10",yTrans)
   yFit     = get(PlotTRY$FitName[p])
   ySumm    = yFit$SummAllom
   
   cat0(" + Plot scatter plot of predicted allometry for ",yDesc,".")

   
   # Copy data object that contains the fitted model.
   PredTRY = yFit$PredAllom

   # Find the number of traits for plotting the model
   CategSel    = is.finite(CategAllom$TraitID)
   CategTrait  = sort(unique(CategAllom$TraitID[CategSel]))
   for (ct in seq_along(CategTrait)){
      # Load settings for the classes.
      zTrait = CategTrait[ct]
      if (zTrait %in% 0L){
         zName  = "Cluster"
         zDesc  = "Data-based cluster class"
      }else{
         z      = match(zTrait,try_trait$TraitID)
         zName  = try_trait$Name   [z]
         zDesc  = try_trait$Desc   [z]
      }#end if (zTrait %in% 0L)
      zTitle = paste0("Allometric model for ",yDesc," by ",tolower(zDesc))

      # Select model and parameters
      zIdx   = which( ( ySumm$TraitClass %in% zName ) & (! is.na(ySumm$Function) ) )[1L]
      zModel = ySumm$Function[zIdx]
      xName  = ySumm$xName   [zIdx]
      wName  = ySumm$wName   [zIdx]

      # Load information for X axis.
      x        = match(xName,try_trait$Name)
      xTraitID = try_trait$TraitID[x]
      xLabel   = sprintf("ID_%+4.4i",xTraitID)
      xLabel   = gsub(pattern="\\+",replacement="p",x=xLabel)
      xLabel   = gsub(pattern="\\-",replacement="m",x=xLabel)
      xDefName = switch( EXPR     = xLabel
                       , ID_m0004 = "LMSize"
                       , ID_m0003 = "WDSize"
                       , ID_m0002 = "Size"
                       , ID_p0018 = "Height"
                       , ID_p0021 = "DBH"
                       , ID_p3106 = "Height"
                       )#end switch
      xName    = try_trait$Name   [x]
      xDesc    = try_trait$Desc   [x]
      xUnit    = try_trait$Unit   [x]
      xTrans   = if(xDefName %in% names(xTransAllom)){xTransAllom[xDefName]}else{"identity"}
      xTPlot   = switch( EXPR = xTrans, log = "log10", neglog="neglog10",xTrans)
      
            
      # Set path for this group of plots.
      categ_allom_path = file.path(allom_path,zName)
      
      # Select categories for this trait type.
      CategNow = CategAllom %>%
         rename( Class = Name) %>%
         filter( (TraitID %in% zTrait) | is.na(TraitID)) %>%
         filter(! Class %in% "UKN") %>%
         filter(! duplicated(Class))
      CntCategNow = nrow(CategNow)

      
      # Find the optimal number of rows and columns, and find multiplication factor 
      # so plots are not ridiculously small
      nColCateg   = min(3,ceiling(sqrt(CntCategNow)))
      nRowCateg   = ceiling(CntCategNow/nColCateg)
      multHeight  = sqrt(nRowCateg)
      multWidth   = multHeight * nColCateg / nRowCateg


      # Select valid data for the axes
      xSel = switch( EXPR     = xTrans
                   , identity = is.finite(PredTRY$x)
                   , log      = PredTRY$x %gt% 0.
                   , neglog   = PredTRY$x %lt% 0.
                   , sqrt     = PredTRY$x %ge% 0.
                   , cbrt     = is.finite(PredTRY$x)
                   , stop(paste0(" Invalid transformation for trait ",xName," (TraitID = ",xTraitID,")."))
                   )#end switch
      ySel = switch( EXPR     = yTrans
                   , identity = is.finite(PredTRY$y)
                   , log      = PredTRY$y %gt% 0.
                   , neglog   = PredTRY$y %lt% 0.
                   , sqrt     = PredTRY$y %ge% 0.
                   , cbrt     = is.finite(PredTRY$y)
                   , stop(paste0(" Invalid transformation for trait ",yName," (TraitID = ",yTraitID,")."))
                   )#end switch

      
      #    Subset data. We keep only valid data points that were either assigned one of the
      # categories of this class, or classified as "All".
      PredPlot    = PredTRY %>%
         filter(xSel & ySel) %>%
         filter( Class %in% CategNow$Class) %>%
         mutate( Class = factor(x=Class,levels=CategNow$Class) )


      # Select valid data for the axes
      xSel = switch( EXPR     = xTrans
                   , identity = is.finite(AllomTRY[[xName]])
                   , log      = AllomTRY[[xName]] %gt% 0.
                   , neglog   = AllomTRY[[xName]] %lt% 0.
                   , sqrt     = AllomTRY[[xName]] %ge% 0.
                   , cbrt     = is.finite(AllomTRY[[xName]])
                   , stop(paste0(" Invalid transformation for trait ",xName," (TraitID = ",xTraitID,")."))
                   )#end switch
      ySel = switch( EXPR     = yTrans
                   , identity = is.finite(AllomTRY[[yName]])
                   , log      = AllomTRY[[yName]] %gt% 0.
                   , neglog   = AllomTRY[[yName]] %lt% 0.
                   , sqrt     = AllomTRY[[yName]] %ge% 0.
                   , cbrt     = is.finite(AllomTRY[[yName]])
                   , stop(paste0(" Invalid transformation for trait ",yName," (TraitID = ",yTraitID,")."))
                   )#end switch
      zSel = AllomTRY[[zName]] %in% CategNow$Class

      AllomShow          = AllomTRY %>% filter(xSel & ySel & zSel)
      AllomShow[[zName]] = factor(x=AllomShow[[zName]],levels=CategNow$Class)
      
      
      #   Make sure all classes are included. In case any class is missing, add a dummy data point with
      # no information.
      TallyClass = table(PredPlot$Class)
      MissClass  = names(TallyClass[TallyClass == 0L])
      if (length(MissClass) > 0L){
           mIdx = match(MissClass,CategNow$Class)

           # Create a single line for missing categories, but leave no data.
           PredMiss = PredPlot[rep(1L,times=length(MissClass)),,drop=FALSE] %>%
              mutate(across(where(is.numeric), ~ .x * NA_real_)) %>%
              mutate( Class      = factor(x=CategNow$Class[mIdx],levels=CategNow$Class)
                    , TraitClass = CategNow$TraitClass[mIdx]
                    , DescClass  = CategNow$DescClass [mIdx]
                    )#end mutate

           # Append dummy data frame
           PredPlot = rbind(PredPlot,PredMiss)
      }#end if (length(MissClass) > 0L)
      
      
      # Decide whether or not to plot the data.
      CntPredPlot = nrow(PredPlot)
      PlotAllom   = CntPredPlot %ge% CntAllomMin
      PlotFit     = any(is.finite(PredPlot$y))

      # Plot only the variables with meaningful data
      if (PlotAllom){
         cat0("   - Plot the allometric model for ",yDesc,", by ",zDesc,".")

         
         # Set path for this group of plots.
         dummy = dir.create(categ_allom_path,recursive=TRUE,showWarnings=FALSE)
 

         # Find limits for plots.   
         xLimit = range(c(PredPlot$x),finite=TRUE)
         yLimit = range(c(PredPlot$y,PredPlot$yLwr,PredPlot$yUpr),finite=TRUE)

         # Set categories for colours, lines and symbols
         colClasses         = CategNow$Colour
         names(colClasses)  = CategNow$Class
         pchClasses         = CategNow$Symbol
         names(pchClasses)  = CategNow$Class
         labClasses         = names(colClasses)
         
         
         # Build the plot
         gg_now = ggplot()
         if (PlotFit){
            gg_now = gg_now + scale_colour_manual(name="",aesthetics="fill"  ,labels=labClasses,values=colClasses)
            gg_now = gg_now + scale_colour_manual(name="",aesthetics="colour",labels=labClasses,values=colClasses)
            gg_now = gg_now + geom_ribbon( data        = PredPlot
                                         , mapping     = aes(x=x,ymin=yLwr,ymax=yUpr,fill=Class)
                                         , alpha       = 0.25
                                         , colour      = "transparent"
                                         , linewidth   = 1.0
                                         , linetype    = "dotdash"
                                         , inherit.aes = FALSE
                                         , show.legend = FALSE
                                         )#end geom_line
            gg_now = gg_now + geom_point( data        = AllomShow
                                        , mapping     = aes_string(x=xName,y=yName,colour=zName)
                                        , size        = 0.5
                                        , shape       = 4L
                                        , inherit.aes = FALSE
                                        , show.legend = TRUE
                                        )#end geom_line
            gg_now = gg_now + geom_line( data        = PredPlot
                                       , mapping     = aes(x=x,y=y,colour=Class)
                                       , linewidth   = 1.5
                                       , linetype    = "solid"
                                       , inherit.aes = FALSE
                                       , show.legend = TRUE
                                       )#end geom_line
            gg_now = gg_now + labs( x        = desc.unit(desc=xDesc,unit=untab[[xUnit]])
                                  , y        = desc.unit(desc=yDesc,unit=untab[[yUnit]])
                                  , colour   = element_blank()
                                  , fill     = element_blank()
                                  , shape    = element_blank()
                                  , linetype = element_blank()
                                  , title    = zTitle
                                  , subtitle = LabelSubtitle
                                  )#end labs
         }else{
            gg_now = gg_now + labs( x        = desc.unit(desc=xDesc,unit=untab[[xUnit]])
                                  , y        = desc.unit(desc=yDesc,unit=untab[[yUnit]])
                                  , colour   = element_blank()
                                  , shape    = element_blank()
                                  , title    = zTitle
                                  , subtitle = LabelSubtitle
                                  )#end labs
         }#end if (PlotFit)
         gg_now = gg_now + theme_grey( base_size = gg_ptsz, base_family = "Helvetica",base_line_size = 0.8,base_rect_size =0.8)
         gg_now = gg_now + theme( axis.text.x       = element_text( size = gg_ptsz, margin = unit(rep(0.35,times=4),"cm"))
                                , axis.text.y       = element_text( size = gg_ptsz, margin = unit(rep(0.35,times=4),"cm"))
                                , plot.title        = element_text( size = gg_ptsz)
                                , plot.subtitle     = element_text( size = 0.7*gg_ptsz)
                                , axis.ticks.length = unit(-0.25,"cm")
                                , legend.position   = "bottom"
                                , legend.direction  = "horizontal"
                                )#end theme
         gg_now = gg_now + guides(fill= guide_colourbar(barwidth=unit(1./6.,"npc")))
         gg_now = gg_now + scale_x_continuous(trans=xTPlot,limits=xLimit)
         gg_now = gg_now + scale_y_continuous(trans=yTPlot,limits=yLimit)

         # Save plot in every format requested.
         for (d in sequence(ndevice)){
            f_output = paste0("ScatterFit_",yName,"_by-",zName,"_",base_suffix,".",gg_device[d])
            dummy    = ggsave( filename = f_output
                             , plot     = gg_now
                             , device   = gg_device[d]
                             , path     = categ_allom_path
                             , width    = gg_square
                             , height   = gg_square
                             , units    = gg_units
                             , dpi      = gg_depth
                             )#end ggsave
         }#end for (o in sequence(nout))
         
         # Write plot settings to the list.
         yzName             = paste0(yName,"_",zName)
         gg_funct[[yzName]] = gg_now
      }else{
         cat0(" + Skip plot for ",yDesc," by ",zDesc,": too few valid points.")
      }#end if (PlotAllom)
   }#end for (ct in seq_along(CategTrait))
}#end for (p in GoodLoop){

# If sought, plot images on screen
cnt_gg_funct = length(gg_funct)
if (gg_screen && (cnt_gg_funct %gt% 0L)){
   gg_show = sort(sample.int(n=cnt_gg_funct,size=min(cnt_gg_funct,3L),replace=FALSE))
   gg_funct[gg_show]
}#end if (gg_screen && (cnt_gg_funct %gt% 0L))
```








# Create a summary table with the allometry

## Goodness-of-fit assessment


In this block, we plot the functional shape of the fitted equations, and compare the results for the different categories. We only plot this if all classes are using the same functional form.

```{r,label='make-summ-allom',message=FALSE,results='hide',warning=FALSE}


# Initialise list of output variables
AllomSumm = NULL
ABootSumm = NULL 

# Loop through all models
cat0(" + Create summary table for all allometric models.")

for (p in sequence(CntPlotTRY)){
   # Retrieve name containing the model fitting object and the trait ID for predictand.
   y        = match(PlotTRY$TraitID[p],try_trait$TraitID)
   yTraitID = try_trait$TraitID[y]
   yDefName = PlotTRY$DefName  [p]
   yName    = try_trait$Name   [y]
   yDesc    = try_trait$Desc   [y]
   yDescObs = paste0(yDesc," - Observed")
   yDescMod = paste0(yDesc," - Fitted"  )
   yUnit    = try_trait$Unit   [y]
   yTrans   = if(yDefName %in% names(yTransAllom)){yTransAllom[yDefName]}else{"identity"}
   yTPlot   = switch( EXPR = yTrans, log = "log10", neglog="neglog10",yTrans)
   yFit     = get(PlotTRY$FitName[p])
   ySumm    = yFit$SummAllom
   yBoot    = yFit$SummBoot
   
   cat0("   - Models for ",yDesc,".")
   AllomSumm = if (is.null(AllomSumm)){ySumm}else{rbind(AllomSumm,ySumm)}
   ABootSumm = if (is.null(ABootSumm)){yBoot}else{rbind(ABootSumm,yBoot)}
}#end for (p in  sequence(CntPlotTRY)){


# Build file name for allometry summary.
allometry_summ  = file.path(summ_path,paste0("TRY_InfoAllom_",base_suffix ,".csv"))
allomboot_summ  = file.path(summ_path,paste0("TRY_BootAllom_",base_suffix ,".csv"))

# Create directory in case it is not available.
dummy = dir.create(summ_path,recursive=TRUE,showWarnings=FALSE)

# Write CSV files with allometry summaries
cat0(" + Write CSV files with allometry model summaries:")
dummy = write_csv( x = AllomSumm, file = allometry_summ, na = "" )
dummy = write_csv( x = ABootSumm, file = allomboot_summ, na = "" )
```
